---
title: "Process expedition taxa"
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_depth: 1
    toc_float: yes
    number_sections: false
    theme: cerulean
    highlight: haddock
---

```{r}
library(raster)
library(sf)
library(ggspatial)
library(lubridate)
library(bdc)
library(tidyverse)
source("_prj_vars.R")
load(file = file.path(ps_exp_path, "data", "secondary", "processed", "admin_spatial_layers.RData"))
```

This scripts creates a master list of all taxa observed during the expedition with their complete taxonomic hierarchy. It also cleans up and corrects scientific names, and pulls relevant data from WORMS, Fish Base, and IUCN. Here we go. 

# Taxa list

## Dropcams

```{r}
dscm_data <- read_csv(file.path(ps_exp_path, 
                                "data/primary/raw",
                                "dscm",
                                "annotations",
                                "COL_dscm_annotations.csv"), 
                      trim_ws = T, skip_empty_rows = T) %>% 
  mutate(taxa = str_to_sentence(taxa))

dscm_meta <- read_csv(file.path(ps_exp_path, "data", "primary", 
                                "processed", "metadata" ,"dscm_meta.csv"))

dscm_data <- dscm_data %>% 
  left_join(dscm_meta %>% select(ps_station_id, location))

dscm_taxa <- dscm_data %>% 
  filter(!str_detect(taxa, "bottom|Plankton"), !is.na(taxa)) %>% 
  distinct(location = location, taxon = taxa) %>% 
  mutate(method = "dscm")
```

```{r}
dscm_data_caribe <- read_csv(file.path(ps_exp_path, 
                                       "data/primary/raw",
                                       "dscm",
                                       "annotations",
                                       "SEA_dropcam_annotations_fish.csv"), 
                             trim_ws = T, skip_empty_rows = T)

dscm_taxa <- dscm_data_caribe %>%
  filter(!is.na(taxa)) %>% 
  left_join(dscm_meta %>% select(ps_station_id, location)) %>% 
  distinct(location = location, taxon = taxa) %>% 
  mutate(method = "dscm") %>% 
  bind_rows(dscm_taxa)
```

## Meso

```{r}
meso_maxN_caribe <- readxl::read_xlsx(file.path(ps_exp_path, 
                                    "data/primary/raw",
                                    "meso","Mesophotic_Caribbean_MaxN.xlsx")) %>% 
  janitor::clean_names()

meso_taxa <- meso_maxN_caribe %>% 
  select(-sharks, -jacks, -snappers, -groupers, -grunts) %>% 
  pivot_longer(!c(location, bruvs_id), names_to = "taxon", values_to = 'maxN', values_drop_na = T) %>% 
  mutate(location = "Caribbean",
         method = "meso",
         taxon = str_to_sentence(str_replace_all(taxon, "_", " "))) %>% 
  distinct(location, method, taxon)
```

## Pelagics

```{r}
pcam_obs_pacific <- readxl::read_xlsx(file.path(ps_exp_path, 
                                    "data/primary/raw",
                                    "pcam","Colombia_2022_Pelagic_Fieldsheet_2022_04_18.xlsx")) %>% 
  janitor::clean_names()

pcam_taxa_pacific <- pcam_obs_pacific %>% 
  filter(location != "Seaflower") %>% 
  mutate(method = "pcam",
         location = if_else(location == "Tribuga", "Golfo de Tribuga", "Offshore Pacific")) %>% 
  distinct(method, location, species_for_summary_report) %>% 
  filter(!is.na(species_for_summary_report)) 

pcam_taxa_pacific <- pcam_taxa_pacific %>% 
  select(-species_for_summary_report) %>% 
  bind_cols(str_split(pcam_taxa_pacific$species_for_summary_report,
                      pattern = ",", simplify = T) %>% 
              as_tibble())

pcam_taxa_pacific <- pcam_taxa_pacific %>% 
  pivot_longer(cols = -c(method, location)) %>% 
  select(-name, method, location, taxa = value) %>% 
  mutate(taxon = ifelse(taxa == "", NA, taxa)) %>% 
  filter(!is.na(taxon)) 

pcam_taxa_pacific <- pcam_taxa_pacific %>% 
  select(method, location) %>% 
  bind_cols(pcam_taxa_pacific %>% 
              pull(taxon) %>% 
              str_extract_all("[A-z]+", simplify = T) %>% 
              as_tibble() %>% 
              unite(taxon, V1:V2, sep = " ", na.rm	= F) %>% 
              mutate(taxon = str_to_sentence(str_trim(str_squish(taxon)))) %>% 
              select(taxon)) %>% 
  distinct(method, location, taxon)
```

```{r}
pcam_taxa_caribe <- readxl::read_xlsx(file.path(ps_exp_path, 
                                    "data/primary/raw",
                                    "pcam", "Colombia_2022_Pelagic_Fieldsheet_Seaflower 2022_05_17.xlsx"), 
                                    sheet = "taxa list") %>% 
  janitor::clean_names() %>% 
  mutate(method = "pcam", 
         location = "Caribbean") %>% 
  distinct(method, location, taxon = scientific_name)

pcam_taxa <- bind_rows(pcam_taxa_caribe, pcam_taxa_pacific)
```

## Birds

```{r}
birds_taxa <- readxl::read_xlsx(file.path(ps_exp_path, 
                                    "data/primary/raw",
                                    "birds","Colombia_Seabirds_Fieldsheet_2022_05_17_lat_longs updated.xlsx"), 
                                    sheet = "Species total") %>% 
  janitor::clean_names() %>% 
  select(binomial, offshore_pacific, tribuga, seaflower)

birds_taxa <- birds_taxa %>% 
  pivot_longer(!binomial) %>% 
  filter(value) %>% 
  select(-value) %>% 
  transmute(method = "birds",
            taxon = binomial,
            location = str_to_title(str_replace_all(name, "_" ," "))) %>% 
  mutate(location = if_else(location == "Tribuga", "Golfo de Tribuga", 
                          if_else(location == "Seaflower", "Caribbean", location))) %>% 
  filter(!is.na(taxon)) %>% 
  distinct()
```

## Bruvs

```{r}
bruvs_maxN <- readxl::read_xlsx(file.path(ps_exp_path, "data/primary/raw/bruvs","COL_bruvs_MaxN.xlsx")) 

bruvs_taxa <- bruvs_maxN %>% 
  mutate(ps_station_id = paste0("COL_bruvs_", 
                                formatC(row_number(), width = 2, flag = 0))) %>% 
  select(-c(1,2,3)) %>% 
  pivot_longer(!ps_station_id, names_to = "taxon", values_to = "max_n") %>% 
  distinct(taxon) %>% 
  mutate(location = "Caribbean",
         method = "bruvs",
         taxon = case_when(taxon == "C. acronotus" ~ "Carcharhinus acronotus",
                           taxon == "C. perezi" ~ "Carcharhinus perezi",
                           taxon == "G. cirratum" ~ "Ginglymostoma cirratum",
                           taxon == "G. cuvier" ~ "Galeocerdo cuvier",
                           taxon == "Rhizoprionodon sp." ~ "Rhizoprionodon",
                           taxon == "S. mokarran" ~ "Sphyrna mokarran",
                           taxon == "N. brevirostris" ~ "Negaprion brevirostris",
                           taxon == "H. americanus" ~ "Hypanus americanus",
                           taxon == "A. narinari" ~ "Aetobatus narinari",
                           taxon == "U. jamaicensis" ~ "Urobatis jamaicensis",
                           taxon == "P. percellens" ~ "Pseudobatos percellens",
                           taxon == "Barracudas" ~ "Barracuda",
                           taxon == "Mackerels" ~ "Scombridae",
                           taxon == "Hogfish" ~ "Bodianus",
                           taxon == "Turtles" ~ "Cheloniidae",
                           taxon == "Dolphins" ~ "Delphinidae",
                           taxon == "Snappers-yellowtails" ~ "Lutjanidae",
                           taxon == "Jacks" ~ "Carangidae",
                           taxon %in% c("Small Groupers", "Large Groupers") ~ "Serranidae",
                           TRUE ~ NA_character_)) %>% 
  filter(!is.na(taxon))
```

## Sub

```{r}
navi_taxa <- readxl::read_xlsx(file.path(ps_exp_path, "data/primary/raw",
                                         "sub", "files/navigator_sub_taxa_lut.xlsx")) %>% 
  janitor::clean_names() %>% 
  mutate(location = "Offshore Pacific",
         method = "sub") %>% 
  distinct(location, method, taxon)

bn_taxa <- readxl::read_xlsx(file.path(ps_exp_path, "data/primary/raw",
                                       "sub","files/bajo_nuevo_annotations.xlsx"),
                             sheet = "fish obs") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(taxa)) %>% 
  distinct(taxon = taxa)

bn_taxa <- bn_taxa %>% 
  add_row(taxon = c("Serranus luciopercatus", "Symphysanodon octoactinus")) %>% 
  mutate(location = "Caribbean",
         method = "sub")
```

## UVS

### Fish

```{r}
ref_fish_list_pacific <- readxl::read_excel(file.path(ps_exp_path, 
                                        "data", "primary","raw","fish","tribuga_fish_species_codes.xlsx")) %>% 
  janitor::clean_names() 

ref_fish_list_caribe <- readxl::read_excel(file.path(ps_exp_path, 
                                        "data", "primary","raw","fish","SEA_fish_species_codes.xlsx")) %>% 
  janitor::clean_names() 

fish_taxa_list <- ref_fish_list_caribe %>%
  select(ps_taxon_code = sp_code_new, 
         taxon = sci_name_new) %>% 
  mutate(location = "Caribbean") %>% 
  bind_rows(
    ref_fish_list_pacific %>%
      select(ps_taxon_code = species, 
             taxon = fish_base_name) %>% 
      mutate(location = "Golfo de Tribuga") ) %>% 
  distinct()

fish_taxa_list$taxon <- str_remove(fish_taxa_list$taxon, "juvenile|species|specues")

exp_fish_obs_pacific <- readxl::read_excel(file.path(ps_exp_path, 
                                        "data", "primary","raw","fish","tribuga_fish_data_final.xlsx")) %>% 
  janitor::clean_names() %>% 
  distinct(species)

exp_fish_obs_caribe <- readxl::read_excel(file.path(ps_exp_path, 
                                        "data", "primary","raw","fish","SEA_fish_final.xlsx")) %>% 
  janitor::clean_names() %>% 
  distinct(species)

exp_fish_obs <- unique(c(exp_fish_obs_caribe$species, exp_fish_obs_pacific$species))

fish_taxa_list <- fish_taxa_list %>% 
  filter(ps_taxon_code %in% exp_fish_obs) %>% 
  mutate(method = "uvs")

fish_taxa_list <- fish_taxa_list %>% 
  mutate(taxon = if_else(stringr::str_count(taxon, boundary("word")) == 1, 
                        paste(taxon, "sp."),
                        taxon)) 
```

### Inverts

```{r}
inverts_taxa_pacific <- readxl::read_excel(file.path(ps_exp_path, 
                                        "data", "primary","raw","inverts","COL_contacts_fieldsheet.xlsx"), 
                               sheet = "All species") %>% 
  janitor::clean_names()

inverts_taxa_caribe <- readxl::read_excel(file.path(ps_exp_path, 
                                        "data", "primary","raw","inverts","SEA_contacts_fieldsheet_definitivo.xlsx"), 
                               sheet = "All species") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(species))

inverts_taxa <- bind_rows(inverts_taxa_caribe %>% 
                            distinct(species, category) %>% 
                            mutate(location = "Caribbean"), 
                          inverts_taxa_pacific %>% 
                            distinct(species, category) %>% 
                            mutate(location = "Golfo de Tribuga"))%>% 
  distinct() %>% 
  mutate(across(!location, str_to_sentence)) %>% 
  mutate(method = "uvc") %>% 
  distinct(method, location, taxon = species)
```

## Master list

```{r}
master_taxa_list <- bind_rows(inverts_taxa, 
                               fish_taxa_list,
                               bn_taxa,
                               navi_taxa,
                               bruvs_taxa,
                               birds_taxa,
                               pcam_taxa,
                               meso_taxa,
                               dscm_taxa) %>%
  mutate(taxon = str_to_sentence(taxon)) %>% 
  distinct() 
```

# Clean up names

```{r}
clean_master_taxa <- bdc::bdc_clean_names(master_taxa_list$taxon) %>% 
  janitor::clean_names() %>% 
  select(taxon = scientific_name, 
         taxon_clean = names_clean) %>% 
  mutate(taxon_clean = coalesce(taxon_clean, 
                                taxadb::clean_names(taxon, lowercase = F))) %>% 
  distinct(taxon, taxon_clean) %>% 
  left_join(master_taxa_list)

clean_master_taxa %>% 
  filter(taxon!= taxon_clean)
```

```{r}
clean_master_taxa <- clean_master_taxa %>% 
  mutate(taxon_clean = case_when(str_detect(taxon, "sponge|Sponge") ~ "Porifera",
                                 str_detect(taxon, "nudibranch") ~ "Nudibranchia",
                                 str_detect(taxon, "bryozoan") ~ "Bryozoan",
                                 taxon == "Pseudoceratina crassa red oval? Foto" ~ "Pseudoceratina",
                                 taxon == "Gonodactylus curacaoensis" ~ "Neogonodactylus curacaoensis",
                                 taxon == "Padina durvillaei" ~ "Padina durvillei",
                                 taxon == "Alosa ruetzeleri" ~ "Ulosa ruetzleri",
                                 taxon == "Ectyospongia ferox" ~ "Ectyoplasia ferox",
                                 taxon == "Chrysoscystis lewisii" ~ "Chrysonephos lewisii",
                                 taxon == "Haliger gigas" ~ "Aliger gigas",
                                 taxon == "Elacanthis rubrifrons" ~ "Elacatinus puncticulatus",
                                 taxon == "Caulerpa verticillata f. charoides" ~ "Caulerpa verticillata f. charoides",
                                 TRUE ~ taxon_clean)) %>% 
    mutate(taxon_clean = str_remove_all(taxon_clean, "decumbent|ruffled|encrusting|black|super|big|Unidentified|unidentified|like|-"),
         taxon_clean = str_remove(taxon_clean, "\\bbis\\b"),
         taxon_clean = str_to_sentence(str_squish(str_squish(str_trim(taxon_clean)))))

clean_master_taxa$taxon_clean[clean_master_taxa$taxon == "Alopias superciliosus"] <- "Alopias superciliosus"

clean_master_taxa %>% 
  filter(taxon!= taxon_clean) 

clean_master_taxa$taxon_clean[clean_master_taxa$taxon == "Barracuda"] <- "Sphyraena barracuda"
clean_master_taxa$taxon_clean[clean_master_taxa$taxon == "Strombus costatus"] <- "Macrostrombus costatus"
clean_master_taxa$taxon_clean[clean_master_taxa$taxon == "Strombus raninus"] <- "Lobatus raninus"
clean_master_taxa$taxon_clean[clean_master_taxa$taxon == "Strombus gallus"] <- "Aliger gallus"
```

```{r}
clean_master_taxa %>% 
  distinct(taxon_clean) %>% 
  arrange(taxon_clean)
```

# Worms match

```{r}
clean_taxa_worms <- clean_master_taxa$taxon_clean %>% 
  split(ceiling(seq_along(clean_master_taxa$taxon_clean)/20)) %>% 
  furrr::future_map_dfr(worrms::wm_records_names) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(scientificname)) %>% 
  mutate(taxon_id = if_else(!is.na(valid_aphia_id),
                              paste0("WORMS:", valid_aphia_id),
                              NA_character_)) %>% 
  select(scientificname, status, taxon_valid_name = valid_name, taxon_rank = rank, taxon_id, 
         taxon_authority = valid_authority,
         kingdom, phylum, class, order, family, genus) %>% 
  distinct() 
```

```{r}
duped_taxa_worms <- clean_taxa_worms %>% 
  janitor::get_dupes(scientificname) 

deduped_taxa_worms <- duped_taxa_worms %>% 
  filter(status %in% c("accepted", "junior subjective synonym")) %>% 
  filter(!(kingdom == "Chromista" & scientificname != "Lobophora")) %>% 
  select(-dupe_count)

clean_taxa_worms <- clean_taxa_worms %>% 
  group_by(scientificname) %>% 
  filter(n() <= 1) %>% # remove duplicates
  bind_rows(deduped_taxa_worms) # add deduped

clean_taxa_worms %>% 
  filter(taxon_valid_name != scientificname)
```

# GBIF match

```{r}
clean_taxa_gbif <- bdc_query_names_taxadb(clean_master_taxa$taxon_clean,
                                         replace_synonyms = TRUE,
                                         suggest_names = TRUE,
                                         suggestion_distance = 0.8,
                                         db = "gbif",
                                         parallel = TRUE,
                                         ncores = 2,
                                         export_accepted = FALSE) %>% 
  janitor::clean_names()

clean_taxa_gbif <- clean_taxa_gbif %>% 
  select(original_search, notes, 
         taxon_valid_name = scientific_name, taxon_rank, taxon_id = accepted_name_usage_id, 
         taxon_authority = scientific_name_authorship, kingdom, phylum, class, order, family, genus) %>% 
  distinct()

clean_taxa_gbif %>% 
  filter(taxon_valid_name != original_search)
# clean_taxa_gbif <- clean_taxa_gbif %>% 
#   filter(!is.na(taxon_id)) %>% 
#   mutate(taxon_id = gsub("^.{0,5}", "", taxon_id))
```

# Diagreements and NAs

```{r}
clean_master_taxa <- clean_master_taxa %>% 
  left_join(clean_taxa_worms %>% 
              select(taxon_clean = scientificname, 
                     worms_sci_name = taxon_valid_name,
                     taxon_rank, 
                     taxon_id,
                     taxon_authority,
                     kingdom, phylum, class, order, family, genus) %>% 
              distinct()) 

clean_master_taxa <- clean_master_taxa %>% 
  coalesce_join(clean_taxa_gbif %>% 
                  select(taxon_clean = original_search, 
                         gbif_sci_name = taxon_valid_name,
                         taxon_rank, 
                         taxon_id,
                         taxon_authority,
                         kingdom, phylum, class, order, family, genus) %>%
                  distinct(), 
                by = "taxon_clean") 

clean_master_taxa %>% 
  filter(worms_sci_name != gbif_sci_name | is.na(worms_sci_name), method == "birds")

clean_master_taxa %>% 
  filter(is.na(worms_sci_name), is.na(gbif_sci_name))

clean_master_taxa <- clean_master_taxa %>% 
  filter(!taxon %in% c("Cca", "Thalobata sp")) 
```

# Merge sources

```{r}
clean_master_taxa <- clean_master_taxa %>% 
  mutate(taxon_sci_name = coalesce(worms_sci_name, gbif_sci_name),
         taxon_rank = str_to_sentence(taxon_rank)) %>% 
  select(-worms_sci_name, -gbif_sci_name, -taxon_clean) %>% 
  distinct() %>% 
  select(method, location, taxon, taxon_sci_name, taxon_rank, taxon_id, ps_taxon_code, everything())
```

# VME

```{r}
all_vme_taxa <- read_csv(file.path(ps_data_path, "VMEs", "files", "all_vme_taxa.csv"))

vme_ind <- read_csv(file.path(ps_data_path, "VMEs", "files", "clean_vme_ind.csv"))

clean_master_taxa <- clean_master_taxa %>% 
  mutate(is_vme = if_else(str_detect(taxon_id, "WORMS:") & 
                            as.numeric(str_remove(taxon_id, "WORMS:")) %in% 
                            c(vme_ind$valid_aphia_id, all_vme_taxa$valid_AphiaID), T, F)) 

clean_master_taxa %>% filter(is_vme)
```

# IUCN

```{r}
iucn_db <- read_csv(file.path(ps_data_path, 
                              "iucn-redlist-marine-species",  
                              "joined_and_resolved_taxa.csv"))
  
iucn_db <- iucn_db %>% 
  filter(taxon_valid_name %in% clean_master_taxa$taxon_sci_name) %>% 
  distinct(taxon_valid_name, iucn_redlist_cat, iucn_trend)  
  
iucn_db %>% 
janitor::get_dupes(taxon_valid_name)

iucn_db <- iucn_db %>% 
  mutate(iucn_redlist_cat = case_when(taxon_valid_name == "Dichocoenia stokesii" ~ "Vulnerable",
                                      taxon_valid_name == "Siderastrea siderea" ~ "Least Concern",
                                      TRUE ~ iucn_redlist_cat),
         iucn_trend = case_when(taxon_valid_name == "Kyphosus vaigiensis" ~ "Unknown",
                                      taxon_valid_name == "Dichocoenia stokesii" ~ "Decreasing",
                                      TRUE ~ iucn_trend)) %>% 
    distinct(taxon_valid_name, iucn_redlist_cat, iucn_trend)  

clean_master_taxa <- clean_master_taxa %>% 
  left_join(iucn_db,
            by = c("taxon_sci_name" = "taxon_valid_name"))

clean_master_taxa %>% 
  group_by(iucn_redlist_cat) %>% 
  summarize(n_distinct(taxon_id))
```

# Fishbase params

```{r}
fb_taxa_names <- clean_master_taxa$taxon_sci_name[clean_master_taxa$taxon_rank == "Species"]

fish_base_estimates <- rfishbase::estimate(fb_taxa_names)

fish_base_estimates <- fish_base_estimates %>% 
  janitor::clean_names() %>% 
  select(fb_taxa_code = spec_code, species, max_length_tl, troph, se_troph, a, sd_log10a, b, sd_b,method_ab, age_min, age_max, feeding_path) %>% 
  distinct()

clean_master_taxa <- clean_master_taxa %>% 
   left_join(fish_base_estimates, 
             by = c("taxon_sci_name"="species")) 

clean_master_taxa %>% 
  filter(is.na(a) & taxon_rank == "Species", 
         class %in% c("Actinopteri", "Actinopterygii", "Elasmobranchii")) %>% 
  distinct()
```

```{r}
fish_missing_lw_params <- tribble(~taxon_sci_name, ~a, ~b, ~troph, ~max_length_tl,
                                  "Carcharhinus perezi", 0.00490, 3.08, 4.5, 300,
                                  "Brachygenys chrysargyreum", 0.01202, 3.00, 3.5, 23)

clean_master_taxa <- clean_master_taxa %>% 
  coalesce_join(fish_missing_lw_params, by = "taxon_sci_name") 
```

```{r}
clean_master_taxa %>% 
  filter(method == "uvc",is.na(a) & class %in% c("Actinopteri", "Actinopterygii", "Elasmobranchii")) %>% 
  distinct(taxon_sci_name)
```

```{r}
impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))

clean_master_taxa <- clean_master_taxa %>% 
  group_by(genus) %>% 
  mutate_at(vars(a, b, troph, max_length_tl), .funs =  impute.mean) %>% 
  mutate_all(~ifelse(is.nan(.), NA, .)) %>% 
  mutate(method_ab = if_else(is.na(method_ab) & !is.na(a), 
                             "Imputed at genus level", 
                             method_ab)) %>% 
  ungroup()
```

# Common names

```{r}
worms_common_names <- clean_master_taxa %>% 
  filter(str_detect(taxon_id, "WORMS")) %>% 
  transmute(worms_id = as.numeric(str_remove_all(taxon_id, "WORMS:"))) %>% 
  pull() %>% 
  worrms::wm_common_id_() %>% 
  filter(language_code %in% c("eng", "spa")) %>% 
  mutate(vernacular = str_squish(str_to_sentence(vernacular))) %>% 
  group_by(id, language_code) %>% 
  summarise(common_names = paste(vernacular, collapse = ";"))

worms_common_names <- worms_common_names %>% 
  pivot_wider(names_from = language_code, 
              names_prefix = "common_name_",
              values_from = common_names)

worms_common_names$id <- paste0("WORMS:", worms_common_names$id)

clean_master_taxa <- clean_master_taxa %>% 
  left_join(worms_common_names, by = c("taxon_id" = "id"))

clean_master_taxa <- clean_master_taxa %>% 
  ungroup() %>% 
  mutate(method = if_else(method == "uvc", "uvs", method)) %>% 
  arrange(taxon_sci_name) 

clean_master_taxa %>%   
write_csv(file.path(ps_exp_path, "data", "primary", "processed","taxa", "clean_master_taxa.csv"))
```

```{r}
clean_master_taxa %>% 
  filter(location == "Caribbean") %>% 
  distinct(taxon_sci_name, taxon_rank, taxon_id, taxon_authority, kingdom, phylum, class, order, family, genus, is_vme, iucn_redlist_cat, iucn_trend) %>% 
  write_csv(file.path(ps_exp_path, "reports/Caribe/Anexos", "Tabla_S2_listado_de_especies.csv"))
```

# New records for the Caribbean

```{r}
c("Chilomycterus antillarum",
  "Amphichthys",
  "Prognathodes guyanensis", 
  "Liopropoma aberrans",
  "Gonioplectrus hispanus",
  "Serranus luciopercanus",
  "Sphyraena borealis",
  "Lutjanus campechanus",
  "Galeus springeri",
  "Alopias superciliosus",
  "Symphysanodon octoactinus") %>% 
  furrr::future_map_dfr(worrms::wm_records_names) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(scientificname)) %>% 
  mutate(taxon_id = if_else(!is.na(valid_aphia_id),
                            paste0("WORMS:", valid_aphia_id),
                            NA_character_)) %>% 
  select(class, order, family, genus, taxon_sci_name = valid_name) %>% 
  distinct() %>% 
  arrange(class, order) %>% 
  write_csv("new_records.csv")
```


