---
title: "Reef fish communities"
output:
  html_document:
    toc: yes
    toc_depth: '1'
    df_print: paged
  word_document:
    toc: yes
    toc_depth: '1'
---

```{r echo = FALSE, include = FALSE, warning = FALSE,prompt = FALSE, error = FALSE, strip.white = TRUE}

knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE,
                      prompt = FALSE, error = FALSE, strip.white = TRUE)

knitr::opts_knit$set(progress = TRUE, verbose = TRUE)

options(dplyr.summarise.inform = FALSE)

library(tidyverse)
library(vegan)

source("_prj_vars.R")

load(file = file.path(ps_exp_path, 
                      "data", "secondary", "processed", "admin_spatial_layers.RData"))


```

```{r}
uvc_meta <- read_csv(file.path(ps_exp_path, 
                               "data", "primary", "processed", "metadata", "uvc_meta.csv"))

uvc_taxa <- read_csv(file.path(ps_exp_path, 
                               "data", "primary", "processed", "taxa", "clean_master_taxa.csv")) %>% 
  filter(method == "uvc")

caribe_fish_data <- readxl::read_xlsx(file.path(ps_exp_path, 
                               "data", "primary","raw", "fish", "SEA_fish_final.xlsx"))
```


```{r}
caribe_fish_data <- caribe_fish_data %>%
  rename(min_length = min, max_length = max, n_ind = number, ps_taxon_code = species) %>%
  rowwise() %>% 
  mutate(avg_length = (min_length + min_length)/2,
         depth_strata = cut(depth_m, breaks = c(0, 15, Inf), labels = c("S", "D"))) %>% 
  select(SEA_station_id = ps_station_id, depth_strata, depth_m, diver, transect, 
         ps_taxon_code, n_ind,avg_length) %>% 
  left_join(uvc_meta %>% 
              filter(location == "Caribbean") %>% 
              distinct(ps_station_id, sublocation, habitat) %>% 
              mutate(SEA_station_id = paste0("SEA_fish_", 
                                             formatC(row_number(), width = 2, flag = 0)))) %>% 
  select(sublocation, ps_station_id, everything(), -SEA_station_id) %>% 
  left_join(uvc_taxa %>% 
              filter(location == "Caribbean") %>% 
              distinct(ps_taxon_code, taxon_sci_name, family, order, class,
                       is_vme, iucn_redlist_cat, max_length_tl, troph, a, b)) %>% 
  mutate(habitat = str_extract(habitat, "[^,]+"))
```

```{r QA/QC}
caribe_fish_data %>%
  filter(avg_length > max_length_tl) %>%
  transmute(abs(avg_length - max_length_tl))

caribe_fish_data <- caribe_fish_data %>% 
  mutate(avg_length = if_else(avg_length > max_length_tl & !is.na(max_length_tl),
                              max_length_tl, 
                              avg_length))
```

```{r}
fish_data_by_transect <- caribe_fish_data %>% 
  mutate(transect_id = paste(ps_station_id, transect, diver, depth_strata, sep = "_"),
         biomass_gr = n_ind*a*avg_length^b,
         ind_m2 = if_else(avg_length <= 20, n_ind/50, n_ind/100),
         gr_m2 = if_else(avg_length <= 20, biomass_gr/50, biomass_gr/100)) %>% 
  group_by(sublocation, ps_station_id, transect_id, transect, diver, depth_strata, depth_m, habitat,
           ps_taxon_code, taxon_sci_name, order, family, iucn_redlist_cat, troph) %>% 
  summarise(across(c(n_ind, ind_m2, biomass_gr, gr_m2), sum)) %>% 
  ungroup() %>% 
  mutate(across(where(is.numeric), round, 3))

write_csv(fish_data_by_transect,
          file.path(ps_exp_path,"data/primary/processed/analyses/uvc_caribe_fish_data_by_transect.csv"))
```

# Taxa composition

We conducted a total of `r n_distinct(caribe_fish_data$ps_station_id)` underwater visuals census to characterize the communities of reef fishes in Bajo Nuevo (n = `r n_distinct(caribe_fish_data$ps_station_id[caribe_fish_data$sublocation == "Bajo Nuevo"])`) and Serranilla (n = `r n_distinct(caribe_fish_data$ps_station_id[caribe_fish_data$sublocation == "Serranilla"])`). In total, we registered `r n_distinct(caribe_fish_data$taxon_sci_name)` distinct taxa, belonging to `r n_distinct(caribe_fish_data$family)` families. 

```{r include = T}
(caribe_fish_diversity_treemap <- caribe_fish_data %>% 
  group_by(sublocation, order, family) %>% 
  summarise(weight = n_distinct(ps_taxon_code))%>% 
  ggplot(aes(area = weight, 
             label = paste(family, weight, sep = "\n"),
             fill = order,
             subgroup = order)) +
  treemapify::geom_treemap(show.legend = T)+
  treemapify::geom_treemap_text(colour = "white", place = "middle", reflow = T, min.size = 3)+
  labs(fill = "",
       title = "Taxa diversity in underwater fish surveys")+
  facet_wrap("sublocation") +
  scale_fill_manual(values = colorRampPalette(RColorBrewer::brewer.pal(12, "Set3"))(24))+
  bbplot::bbc_style()+
  guides(fill = guide_legend(nrow=5,byrow=TRUE))+
  theme(legend.position = "bottom"))

ggsave(caribe_fish_diversity_treemap, 
       filename = file.path(ps_exp_path, "figures", "uvc_caribe_fish_diversity_treemap.png"), 
       dpi = 300, width = 12, height = 10)
```

```{r}
caribe_fish_data %>% 
  group_by(sublocation) %>% 
  summarise(n_stations = n_distinct(ps_station_id),
            n_families = n_distinct(family),
            n_species = n_distinct(ps_taxon_code),
            n_ind = sum(n_ind))
```
```{r}
caribe_fish_abundance_treemap <- fish_data_by_transect %>% 
  group_by(sublocation, ps_station_id, ps_taxon_code, order, family) %>% 
  summarise(across(c(ind_m2, gr_m2), mean)) %>%
  ungroup() %>% 
  group_by(sublocation, order, family) %>% 
  summarise(across(c(ind_m2, gr_m2), sum)) %>% 
  ungroup() %>% 
  mutate(f_ind_m2 = round(100*ind_m2/sum(ind_m2),1),
         f_gr_m2 = round(100*gr_m2/sum(gr_m2), 1)) %>% 
  ggplot(aes(area = ind_m2, 
             label = paste(family, paste0(f_ind_m2, "%"), sep = "\n"),
             fill = order,
             subgroup = order)) +
  treemapify::geom_treemap(show.legend = T)+
  treemapify::geom_treemap_text(colour = "white", place = "middle", reflow = T, min.size = 3)+
  labs(fill = "",
       title = "Taxa abundance (ind m2) in underwater fish surveys")+
  facet_wrap("sublocation", nrow = 2) +
  scale_fill_manual(values = colorRampPalette(RColorBrewer::brewer.pal(12, "Set3"))(24))+
  #bbplot::bbc_style()+
  theme(legend.position = "bottom")
  #guides(fill = guide_legend(nrow=5,byrow=TRUE))+
  # theme(legend.position = "rigth")

ggsave(caribe_fish_abundance_treemap, 
       filename = file.path(ps_exp_path, "figures", "uvc_caribe_fish_abundance_treemap.png"),
       dpi = 300, width = 8, height = 10)
```

```{r}
caribe_fish_biomass_treemap <- fish_data_by_transect %>% 
  group_by(sublocation, ps_station_id, ps_taxon_code, order, family) %>% 
  summarise(across(c(ind_m2, gr_m2), mean)) %>%
  ungroup() %>% 
  group_by(sublocation, order, family) %>% 
  summarise(across(c(ind_m2, gr_m2), sum)) %>% 
  ungroup() %>% 
  mutate(f_ind_m2 = round(100*ind_m2/sum(ind_m2),1),
         f_gr_m2 = round(100*gr_m2/sum(gr_m2), 1)) %>% 
  ggplot(aes(area = gr_m2, 
             label = paste(family, paste0(f_gr_m2, "%"), sep = "\n"),
             fill = order,
             subgroup = order)) +
  treemapify::geom_treemap(show.legend = T)+
  treemapify::geom_treemap_text(colour = "white", place = "middle", reflow = T, min.size = 3)+
  labs(fill = "",
       title = "Taxa biomass (gr m2) in underwater fish surveys")+
  facet_wrap("sublocation", nrow = 2) +
  scale_fill_manual(values = colorRampPalette(RColorBrewer::brewer.pal(12, "Set3"))(24))+
  #bbplot::bbc_style()+
  theme(legend.position = "bottom")
  #guides(fill = guide_legend(nrow=5,byrow=TRUE))+
  # theme(legend.position = "rigth")

ggsave(caribe_fish_biomass_treemap, 
       filename = file.path(ps_exp_path, "figures", "uvc_caribe_fish_biomass_treemap.png"),
       dpi = 300, width = 8, height = 10)
```

# Visualize data

```{r}
n_ind_data_wide <- fish_data_by_transect %>% 
  select(transect_id, ps_taxon_code, n_ind) %>% 
  pivot_wider(names_from = ps_taxon_code, values_from = n_ind, values_fill = 0) %>% 
  column_to_rownames("transect_id") 

diversity_by_transect <- diversity(n_ind_data_wide) %>% 
  enframe() %>% 
  set_names(c("transect_id", "H")) %>% 
  mutate(richness = specnumber(n_ind_data_wide),
         evenness = H/log(richness))

transect_summary <- fish_data_by_transect %>% 
  group_by(ps_station_id, transect_id, habitat, depth_strata, depth_m, diver, sublocation) %>% 
  summarize(n_ind = sum(n_ind),
            ind_m2 = sum(ind_m2),
            gr_m2 = sum(gr_m2)) %>% 
  ungroup() %>% 
  left_join(diversity_by_transect)

transect_summary %>% 
  pivot_longer(c(n_ind, ind_m2, gr_m2, H, richness, evenness)) %>% 
  ggplot()+
  geom_histogram(aes(value))+
  facet_wrap("name", scales = "free")
```

```{r}
transect_summary %>% 
  pivot_longer(c(ind_m2, gr_m2, H, richness, evenness)) %>% 
  ggplot()+
  geom_boxplot(aes(x = sublocation, y = value, fill = diver))+
  facet_wrap("name", scales = "free")

transect_summary %>% 
  pivot_longer(c(ind_m2, gr_m2, H, richness, evenness)) %>% 
  ggplot()+
  geom_boxplot(aes(x = sublocation, y = value, fill = habitat))+
  facet_wrap("name", scales = "free")

transect_summary %>% 
  pivot_longer(c(ind_m2, gr_m2, H, richness, evenness)) %>% 
  ggplot()+
  geom_boxplot(aes(x = sublocation, y = value, fill = depth_strata))+
  facet_wrap("name", scales = "free")
```

## Alpha diversity

```{r}
# Data are not normally distributed
transect_summary %>% 
  select(ind_m2, gr_m2, H, evenness) %>% 
  rstatix::shapiro_test(ind_m2, gr_m2, H, evenness)
```

```{r}
adonis2(transect_summary %>% 
          select(ind_m2) ~  diver + depth_strata + habitat + sublocation, 
        data = transect_summary,
        strata = transect_summary$ps_station_id,
        by = "margin",
        method = "euclidean")

adonis2(transect_summary %>% 
          select(gr_m2) ~  diver + depth_strata + habitat + sublocation, 
        data = transect_summary,
        strata = transect_summary$ps_station_id,
        by = "margin",
        method = "euclidean")

adonis2(transect_summary %>% 
          select(H) ~  diver + depth_strata + habitat + sublocation, 
        data = transect_summary,
        strata = transect_summary$ps_station_id,
        by = "margin",
        method = "euclidean")

adonis2(transect_summary %>% 
          select(H, gr_m2, ind_m2) ~  diver + depth_strata + habitat + sublocation, 
        data = transect_summary,
        strata = transect_summary$ps_station_id,
        by = "margin")
```
```{r}
## Q: Should Shannon index be computed per transect and then averaged ?
caribe_fish_diversity_barplot <- transect_summary %>% 
  group_by(sublocation, habitat) %>% 
  summarize(across(c(H, evenness), list(avg = mean, sd = sd), na.rm = T)) %>% 
  pivot_longer(-c(sublocation, habitat), 
               names_to = c('variable', 'metric'),
               names_sep = "_",
               values_to = "score") %>% 
  pivot_wider(names_from = metric, values_from = score) %>% 
  ggplot()+
  geom_col(aes(sublocation, avg, fill = fct_reorder(habitat, avg)), 
           position = position_dodge2(width = 0.9, preserve = "single"),
           alpha = 0.9) +
  geom_errorbar(aes(x = sublocation,
                    ymax = avg + sd,
                    ymin = avg - sd,
                    fill = fct_reorder(habitat, avg)), 
                position = position_dodge2(width = 0.1, preserve = "single"),
                alpha = 0.5)+
  #coord_flip()+
  paletteer::scale_fill_paletteer_d("ggsci::default_jama")+
  theme(legend.position = "bottom", plot.title.position = "plot") +
  guides(fill = guide_legend(nrow=2,byrow=TRUE))+
  labs(fill = "", x = "", title = "Reef fish species diversity")+
  facet_wrap(~variable, scales = "free")
  
ggsave(caribe_fish_diversity_barplot, 
       filename = file.path(ps_exp_path, "figures", "uvc_caribe_fish_diversity_barplot.png"),       
       dpi = 300, width = 6, height = 4)
```


## Community dissimilarity

```{r}
min_ok_n <- 10

max_n_samp <- min(transect_summary$n_ind[transect_summary$n_ind > min_ok_n])

n_ind_data_wide_trimmed <- n_ind_data_wide[transect_summary$transect_id[transect_summary$n_ind > min_ok_n],]

cols_to_rm <- colSums(n_ind_data_wide[row.names(n_ind_data_wide_trimmed),]) == 0
       
n_ind_data_wide_trimmed <- n_ind_data_wide_trimmed[,!cols_to_rm] 

set.seed(19900623)

n_ind_dist <- avgdist(as.matrix(n_ind_data_wide_trimmed), 
                      dmethod = "bray", 
                      sample = max_n_samp) 
```

### PCOA

#### Unconstrained

```{r}
fish_pcoa <- dbrda(n_ind_dist ~ 1, 
                   data = transect_summary[transect_summary$n_ind > min_ok_n, ],
                   add = "lingoes")

fish_pcoa_env_fit <- envfit(fish_pcoa, 
                            transect_summary %>% 
                              filter(n_ind > min_ok_n) %>% 
                              select(habitat, depth_strata, diver, sublocation))

fish_pcoa_spp_fit <- envfit(fish_pcoa, 
                            n_ind_data_wide_trimmed)

fish_pcoa_spp_scores <- fish_pcoa_spp_fit %>% 
  scores("vectors") %>% 
  as_tibble(rownames = "ps_taxon_code") %>% 
  mutate(r = fish_pcoa_spp_fit$vectors$r,
         p = fish_pcoa_spp_fit$vectors$pvals) %>% 
  filter(p < 0.05, r > 0.1)

fish_pcoa_axis <- BiodiversityR::axis.long(fish_pcoa, choices = c(1, 2))

fish_pcoa_site_scores <- scores(fish_pcoa) %>% 
  as_tibble(rownames = "transect_id") %>% 
  inner_join(transect_summary)
```

```{r}
map(c("habitat", "sublocation", "diver", "depth_strata"), 
    ~ ggplot() +
      geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
      geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
      scale_x_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      geom_point(data = fish_pcoa_site_scores,
                 aes(x = MDS1, y = MDS2,, col = .data[[.x]])) +
      labs(x = fish_pcoa_axis$label[1], y = fish_pcoa_axis$label[2])+
      geom_segment(data = fish_pcoa_spp_scores %>% 
                     filter(p < 0.05, r > 0.2), 
                   aes(x = 0, y = 0, xend = MDS1*3, yend = MDS2*3, alpha = r), 
                   colour = "black", 
                   size = 0.2, 
                   arrow = arrow(length = unit(0.01, "npc"), 
                                 type = 'open', ends = "last"))+
      ggrepel::geom_text_repel(data = fish_pcoa_spp_scores %>% 
                                 filter(p < 0.05, r > 0.2), 
                               aes(x=MDS1*3, y=MDS2*3, label = ps_taxon_code, alpha = r),
                               colour="black",
                               show.legend = F) +
      coord_fixed(ratio=1)+
      ggforce::geom_mark_ellipse(data = fish_pcoa_site_scores, 
                                 aes(x=MDS1, y = MDS2, colour = .data[[.x]], 
                                     fill=after_scale(alpha(colour, 0.01))), 
                                 expand=0, size=0.2, show.legend=FALSE)+
      theme(panel.background = element_blank(),
            panel.border = element_blank(),
            panel.grid = element_blank(),
            axis.line = element_line("gray25"),
            text = element_text(size = 12, family = "Arial"),
            axis.text = element_text(size = 10, colour = "gray25"),
            axis.title = element_text(size = 12, colour = "gray25"),
            legend.key = element_blank())+
      scale_alpha(range=c(0.6,1))
)
```

```{r}
fish_pcoa_env_fit
```

#### Constrained

```{r}
fish_pcoa_c <- dbrda(n_ind_dist ~ diver + habitat + sublocation + depth_strata, 
                    data = transect_summary[transect_summary$n_ind > min_ok_n, ],
                    dist = "bray", 
                    add = "lingoes")

fish_pcoa_c_env_fit <- envfit(fish_pcoa_c, 
                            transect_summary %>% 
                              filter(n_ind > min_ok_n) %>% 
                              select(habitat, depth_strata, diver, sublocation))

fish_pcoa_c_spp_fit <- envfit(fish_pcoa_c, 
                            n_ind_data_wide_trimmed)

fish_pcoa_c_spp_scores <- fish_pcoa_c_spp_fit %>% 
  scores("vectors") %>% 
  as_tibble(rownames = "ps_taxon_code") %>% 
  mutate(r = fish_pcoa_c_spp_fit$vectors$r,
         p = fish_pcoa_c_spp_fit$vectors$pvals) %>% 
  filter(p < 0.05, r > 0.1)

fish_pcoa_c_axis <- BiodiversityR::axis.long(fish_pcoa_c, choices = c(1, 2))

fish_pcoa_c_site_scores <- scores(fish_pcoa_c)$sites %>% 
  as_tibble(rownames = "transect_id") %>% 
  inner_join(transect_summary)

#summary(fish_pcoa_c)
```

```{r}
map(c("habitat", "sublocation", "diver", "depth_strata"), 
    ~ ggplot() +
      geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
      geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
      scale_x_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      geom_point(data = fish_pcoa_c_site_scores,
                 aes(x = dbRDA1, y = dbRDA2,, col = .data[[.x]])) +
      labs(x = fish_pcoa_c_axis$label[1], y = fish_pcoa_c_axis$label[2])+
      geom_segment(data = fish_pcoa_c_spp_scores %>% 
                     filter(p < 0.05, r > 0.2), 
                   aes(x = 0, y = 0, xend = dbRDA1*5, yend = dbRDA2*5, alpha = r), 
                   colour = "black", 
                   size = 0.2, 
                   arrow = arrow(length = unit(0.01, "npc"), 
                                 type = 'open', ends = "last"))+
      ggrepel::geom_text_repel(data = fish_pcoa_c_spp_scores %>% 
                                 filter(p < 0.05, r > 0.2), 
                               aes(x=dbRDA1*5, y=dbRDA2*5, label = ps_taxon_code, alpha = r),
                               colour="black",
                               show.legend = F) +
      coord_fixed(ratio=1)+
      ggforce::geom_mark_ellipse(data = fish_pcoa_c_site_scores, 
                                 aes(x=dbRDA1, y = dbRDA2, colour = .data[[.x]], 
                                     fill=after_scale(alpha(colour, 0.01))), 
                                 expand=0, size=0.2, show.legend=FALSE)+
      theme(panel.background = element_blank(),
            panel.border = element_blank(),
            panel.grid = element_blank(),
            axis.line = element_line("gray25"),
            text = element_text(size = 12, family = "Arial"),
            axis.text = element_text(size = 10, colour = "gray25"),
            axis.title = element_text(size = 12, colour = "gray25"),
            legend.key = element_blank())+
      scale_alpha(range=c(0.6,1))
)
```

```{r}
anova(fish_pcoa_c) # overall test of the significant of the analysis
anova(fish_pcoa_c, by = "axis", perm.max = 500) # test axes for significance
anova(fish_pcoa_c, by = "terms", perm.max = 10000) # test for sign. environ. variables
```

### NMDS

```{r}
n_ind_MDS <- metaMDS(n_ind_dist, k = 2, trymax = 1000)

n_ind_MDS$stress

stressplot(n_ind_MDS)

n_ind_MDS_env_fit <- envfit(n_ind_MDS, 
                            transect_summary %>% 
                              filter(n_ind > min_ok_n) %>% 
                              select(habitat, depth_strata, diver, sublocation))

n_ind_MDS_spp_fit <- envfit(n_ind_MDS, 
                            n_ind_data_wide_trimmed)

n_ind_MDS_spp_scores <- n_ind_MDS_spp_fit %>% 
  scores("vectors") %>% 
  as_tibble(rownames = "ps_taxon_code") %>% 
  mutate(r = n_ind_MDS_spp_fit$vectors$r,
         p = n_ind_MDS_spp_fit$vectors$pvals) %>% 
  filter(p < 0.05, r > 0.1)

n_ind_MDS_axis <- BiodiversityR::axis.long(n_ind_MDS, choices = c(1, 2))

n_ind_MDS_site_scores <- scores(n_ind_MDS, display = "sites") %>% 
  as_tibble(rownames = "transect_id") %>% 
  inner_join(transect_summary)
```

```{r}
map(c("habitat", "sublocation", "diver", "depth_strata"), 
    ~ ggplot() +
      geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
      geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
      scale_x_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      geom_point(data = n_ind_MDS_site_scores,
                 aes(x = NMDS1, y = NMDS2,, col = .data[[.x]])) +
      labs(x = n_ind_MDS_axis$label[1], y = n_ind_MDS_axis$label[2])+
      geom_segment(data = n_ind_MDS_spp_scores %>% 
                     filter(p < 0.05, r > 0.2), 
                   aes(x = 0, y = 0, xend = NMDS1*1, yend = NMDS2*1, alpha = r), 
                   colour = "black", 
                   size = 0.2, 
                   arrow = arrow(length = unit(0.01, "npc"), 
                                 type = 'open', ends = "last"))+
      ggrepel::geom_text_repel(data = n_ind_MDS_spp_scores %>% 
                                 filter(p < 0.05, r > 0.2), 
                               aes(x=NMDS1*1, y=NMDS2*1, label = ps_taxon_code, alpha = r),
                               colour="black",
                               show.legend = F) +
      coord_fixed(ratio=1)+
      ggforce::geom_mark_ellipse(data = n_ind_MDS_site_scores, 
                                 aes(x=NMDS1, y = NMDS2, colour = .data[[.x]], 
                                     fill=after_scale(alpha(colour, 0.01))), 
                                 expand=0, size=0.2, show.legend=FALSE)+
      theme(panel.background = element_blank(),
            panel.border = element_blank(),
            panel.grid = element_blank(),
            axis.line = element_line("gray25"),
            text = element_text(size = 12, family = "Arial"),
            axis.text = element_text(size = 10, colour = "gray25"),
            axis.title = element_text(size = 12, colour = "gray25"),
            legend.key = element_blank())+
      scale_alpha(range=c(0.6,1))
)
```

```{r}
n_ind_adonis <- adonis2(n_ind_dist ~ diver + habitat + depth_strata + sublocation,
        data = transect_summary[transect_summary$n_ind > 10, ],
        strata = transect_summary[transect_summary$n_ind > 10, ]$ps_station_id,
        by = "margin")

n_ind_adonis
```

```{r}
sppscores(n_ind_MDS) <- n_ind_data_wide_trimmed

n_ind_MDS_spp_scores_2 <- scores(n_ind_MDS)$species %>% 
  as_tibble(rownames = "ps_taxon_code") 

map(c("habitat", "sublocation", "diver", "depth_strata"), 
    ~ ggplot() +
      geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
      geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
      scale_x_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      geom_point(data = n_ind_MDS_site_scores,
                 aes(x = NMDS1, y = NMDS2,, col = .data[[.x]])) +
      labs(x = n_ind_MDS_axis$label[1], y = n_ind_MDS_axis$label[2])+
      geom_segment(data = n_ind_MDS_spp_scores_2 , 
                   aes(x = 0, y = 0, xend = NMDS1*.5, yend = NMDS2*.5), 
                   colour = "black", 
                   size = 0.2, 
                   arrow = arrow(length = unit(0.01, "npc"), 
                                 type = 'open', ends = "last"))+
      ggrepel::geom_text_repel(data = n_ind_MDS_spp_scores_2, 
                               aes(x=NMDS1*.5, y=NMDS2*.5, label = ps_taxon_code),
                               colour="black",
                               show.legend = F) +
      coord_fixed(ratio=1)+
      ggforce::geom_mark_ellipse(data = n_ind_MDS_site_scores, 
                                 aes(x=NMDS1, y = NMDS2, colour = .data[[.x]], 
                                     fill=after_scale(alpha(colour, 0.01))), 
                                 expand=0, size=0.2, show.legend=FALSE)+
      theme(panel.background = element_blank(),
            panel.border = element_blank(),
            panel.grid = element_blank(),
            axis.line = element_line("gray25"),
            text = element_text(size = 12, family = "Arial"),
            axis.text = element_text(size = 10, colour = "gray25"),
            axis.title = element_text(size = 12, colour = "gray25"),
            legend.key = element_blank())+
      scale_alpha(range=c(0.6,1))
)

```

## Summary tables

```{r}
fish_data_by_transect %>% 
  group_by(sublocation) %>% 
  summarize(n_stations = n_distinct(ps_station_id),
            n_transects = n_distinct(transect_id),
            n_taxa = n_distinct(ps_taxon_code)) %>% 
  left_join(transect_summary %>% 
              group_by(sublocation) %>% 
               summarize(across(c(ind_m2, gr_m2, H), list(avg = mean, sd = sd)))) %>% 
    mutate_if(is.numeric, round, 2)
```

```{r}
fish_data_by_transect %>% 
  group_by(sublocation, taxon_sci_name) %>% 
  summarise(n_transects = n_distinct(transect_id)) %>% 
  ungroup() %>% 
  group_by(sublocation) %>% 
  mutate(f_obs = n_transects/n_distinct(n_transects)) %>% 
  ungroup() %>% 
  left_join(fish_data_by_transect %>% 
              group_by(sublocation, taxon_sci_name) %>% 
              summarize(across(c(ind_m2, gr_m2), list(avg = mean, sd = sd))) %>% 
              ungroup()) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  arrange(desc(gr_m2_avg))
```

