---
title: "Reef fish communities"
output:
  word_document:
    toc: yes
    toc_depth: '1'
  html_document:
    toc: yes
    toc_depth: '1'
    df_print: paged
---

```{r echo = FALSE, include = FALSE, warning = FALSE,prompt = FALSE, error = FALSE, strip.white = TRUE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE,
                      prompt = FALSE, error = FALSE, strip.white = TRUE)

knitr::opts_knit$set(progress = TRUE, verbose = TRUE)

options(dplyr.summarise.inform = FALSE)

library(sf)
library(tidyverse)
library(vegan)

source("_prj_vars.R")

load(file = file.path(ps_exp_path, 
                      "data", "secondary", "processed", "admin_spatial_layers.RData"))

uvc_meta <- read_csv(file.path(ps_exp_path, 
                               "data", "primary", "processed", "metadata", "uvs_meta.csv")) %>% 
  filter(method == "Fish survey", location == "Caribbean") %>% 
  mutate(habitat = str_extract(habitat, "[^,]+"))

uvc_taxa <- read_csv(file.path(ps_exp_path, 
                               "data", "primary", "processed", "taxa", "clean_master_taxa.csv"))%>% 
  filter(method == "uvs")

caribe_fish_data <- readxl::read_xlsx(file.path(ps_exp_path, 
                               "data", "primary","raw", "fish", "SEA_fish_final.xlsx"))
```

```{r}
uvc_meta %>% 
  group_by(sublocation, habitat) %>% 
  summarize(n_distinct(ps_station_id))
```

```{r}
caribe_sat_img <- terra::rast(file.path(ps_exp_path, 
                               "data", "secondary", "raw", 
                               "Allen atlas", "Coral-Reefs-2020-Visual-V1-Mosaic", "satellite_imagery_0.tif")) 

bajos_eez <- bajos_eez %>% 
  st_transform(terra::crs(caribe_sat_img)) 
  
bn_sat_img <- caribe_sat_img %>% 
  terra::crop(bajos_eez %>% 
                filter(territory1 == "Bajo Nuevo Bank") %>% 
                terra::ext(), mask = T) 

serr_sat_img <- caribe_sat_img %>% 
  terra::crop(bajos_eez %>% 
                filter(territory1 == "Serranilla Bank") %>% 
                terra::ext(), mask = T) 
```

# Clean data 

```{r}
caribe_fish_data <- caribe_fish_data %>%
  rename(min_length = min, max_length = max, n_ind = number, ps_taxon_code = species) %>%
  rowwise() %>% 
  mutate(avg_length = (min_length + min_length)/2,
         depth_strata = cut(depth_m, breaks = c(0, 15, Inf), labels = c("S", "D"))) %>% 
  dplyr::select(SEA_station_id = ps_station_id, depth_strata, depth_m, diver, transect, 
         ps_taxon_code, n_ind,avg_length) %>% 
  left_join(uvc_meta %>% 
              filter(location == "Caribbean") %>% 
              distinct(ps_station_id, sublocation, habitat) %>% 
              mutate(SEA_station_id = paste0("SEA_fish_", 
                                             formatC(row_number(), width = 2, flag = 0)))) %>% 
  dplyr::select(sublocation, ps_station_id, everything(), -SEA_station_id) %>% 
  left_join(uvc_taxa %>% 
              filter(location == "Caribbean") %>% 
              distinct(ps_taxon_code, taxon_sci_name, family, order, class,
                       is_vme, iucn_redlist_cat, max_length_tl, troph, a, b)) 
## QA/QC

caribe_fish_data %>%
  filter(avg_length > max_length_tl) %>%
  transmute(abs(avg_length - max_length_tl))

caribe_fish_data <- caribe_fish_data %>% 
  mutate(avg_length = if_else(avg_length > max_length_tl & !is.na(max_length_tl),
                              max_length_tl, 
                              avg_length))
```

## By transect 

```{r}
fish_data_by_transect <- caribe_fish_data %>% 
  mutate(transect_id = paste(ps_station_id, transect, diver, depth_strata, sep = "_"),
         biomass_gr = n_ind*a*avg_length^b,
         ind_m2 = if_else(avg_length <= 20, n_ind/50, n_ind/100),
         gr_m2 = if_else(avg_length <= 20, biomass_gr/50, biomass_gr/100)) %>% 
  group_by(sublocation, ps_station_id, transect_id, transect, diver, depth_strata, depth_m, habitat,
           ps_taxon_code, taxon_sci_name, order, family, iucn_redlist_cat, troph) %>% 
  summarise(across(c(n_ind, ind_m2, biomass_gr, gr_m2), sum)) %>% 
  ungroup() 

# Need to add zeros

fish_data_by_transect <- fish_data_by_transect %>% 
  distinct(sublocation, ps_station_id, transect_id, diver, depth_strata, depth_m, habitat) %>% 
  full_join(expand.grid(transect_id = unique(fish_data_by_transect$transect_id),
                        ps_taxon_code = unique(fish_data_by_transect$ps_taxon_code))) %>% 
  left_join(fish_data_by_transect %>% 
              distinct(transect_id, ps_taxon_code, n_ind, ind_m2, gr_m2)) %>% 
  left_join(fish_data_by_transect %>% 
              distinct(ps_taxon_code, taxon_sci_name, order, family, iucn_redlist_cat)) %>% 
  replace_na(list(n_ind = 0, ind_m2 = 0, gr_m2 = 0))
```


```{r}
transect_summary <- fish_data_by_transect %>% 
  group_by(sublocation, ps_station_id, transect_id, habitat, depth_strata, depth_m, diver) %>% 
  summarize(n_taxa = n_distinct(ps_taxon_code[n_ind > 0]),
            n_ind = sum(n_ind),
            ind_m2 = sum(ind_m2),
            gr_m2 = sum(gr_m2)) %>% 
  ungroup() 

n_ind_data_wide <- fish_data_by_transect %>% 
  dplyr::select(transect_id, ps_taxon_code, n_ind) %>% 
  pivot_wider(names_from = ps_taxon_code, values_from = n_ind, values_fill = 0) %>% 
  column_to_rownames("transect_id") 

diversity_by_transect <- diversity(n_ind_data_wide) %>% 
  enframe() %>% 
  set_names(c("transect_id", "H")) %>% 
  mutate(richness = specnumber(n_ind_data_wide),
         evenness = H/log(richness))

transect_summary <- transect_summary %>% 
  left_join(diversity_by_transect) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  ungroup() 
```

## By station

```{r}
fish_data_by_station <- fish_data_by_transect %>% 
  group_by(sublocation, habitat, ps_station_id, family, taxon_sci_name) %>% 
  summarize(n_ind = mean(n_ind, na.rm = T),
            ind_m2 = mean(ind_m2, na.rm = T),
            gr_m2 = mean(gr_m2, na.rm = T)) %>% 
  ungroup() %>% 
  group_by(ps_station_id) %>% 
  mutate(f_n_ind = n_ind/sum(n_ind),
         f_ind_m2 = ind_m2/sum(ind_m2),
         f_gr_m2 = gr_m2/sum(gr_m2)) %>% 
  ungroup() %>% 
  mutate_if(is.numeric, round, 2) %>% 
  dplyr::select(sublocation, habitat, ps_station_id, family, taxon_sci_name, 
         n_ind, f_n_ind, ind_m2, f_ind_m2, gr_m2, f_gr_m2)

fish_data_by_station %>% 
  write_csv(file.path(ps_exp_path, "reports/Caribe/Anexos", "Tabla_S3_abundancia_y_biomasa_de_peces_por_taxon_por_estacion.csv"))

station_summary <- transect_summary %>% 
  group_by(sublocation, habitat, ps_station_id) %>% 
  summarize(across(c(n_taxa, n_ind, ind_m2, gr_m2, H, evenness), mean)) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  ungroup() 

station_summary %>% 
  write_csv(file.path(ps_exp_path, "reports/Caribe/Anexos", "Tabla_S4_metricas_univariadas_de_peces_por_estacion.csv"))

```

```{r}
station_summary <- read_csv(file.path(ps_exp_path, 
                                      "reports/Caribe/Anexos",
                                    "Tabla_S4_metricas_univariadas_de_peces_por_estacion.csv"))
```

```{r, eval = F}
serr_map <- station_summary %>% 
  left_join(uvc_meta %>% 
              distinct(ps_station_id, lat, lon)) %>% 
  filter(sublocation == "Serranilla") %>% 
  sf::st_as_sf(coords = c("lon", "lat"), crs = 4326) %>% 
  ggplot()+
  tidyterra::geom_spatraster_rgb(data = serr_sat_img, maxcell = 10^7)+
  geom_sf(aes(col = habitat, size = gr_m2))+
  coord_sf(xlim = c(-79.86, -79.82), ylim = c(15.785, 15.825))+
  ggspatial::annotation_scale(location = "bl", 
                              width_hint = 0.2, 
                              style = "ticks", 
                              line_col = "white",
                              text_col = "white",
                              pad_y = unit(0.01, units =  "native"), 
                              pad_x = unit(0.01, units =  "native"))+
  ggspatial::annotation_north_arrow(location = "tr", 
                                    which_north = "true", 
                                    height = unit(0.1, "native"), width = unit(0.1, "native"),
                                    pad_x = unit(0.01, "native"), pad_y = unit(0.01, "native"),
                                    style = ggspatial::north_arrow_fancy_orienteering)+
  paletteer::scale_color_paletteer_d("RColorBrewer::Set3")+
  scale_shape_manual(values= seq(100:110))+
  theme_light()+
  labs(col = "Habitat", size = bquote('Biomass density '(g/m^2)))+
  guides(size=guide_legend(override.aes=list(colour="grey")))+
  theme(legend.background = element_blank(),
        legend.position = "right",
        #legend.box = 'vertical',
        #legend.justification = 'left',
        #legend.box.just = 'left',
        legend.key = element_blank(),
        legend.text = element_text(colour = "black"), 
        panel.background  = element_rect(fill ="#1b2b49"),
        panel.grid = element_blank())

serr_map_spanish <- serr_map +
    PristineSeasR::scale_color_pristine_seas(palette = "alternative",
                                             labels = c("Terraza prearrecifal", "Arrecife periferico","Cuenca lagunar", "Pastos marinos"))+
    labs(col = "Habitat", size = bquote('Biomasa '(g/m^2)))

ggsave(serr_map_spanish,
       filename = file.path(ps_exp_path, "figures", "serranilla_fish_biomass_map_spanish.png"), 
       width = 8, height = 6,
       dpi = 300)
```

```{r, eval = F}
bn_map <- station_summary %>% 
  left_join(uvc_meta %>% 
              distinct(ps_station_id, lat, lon)) %>% 
  filter(sublocation == "Bajo Nuevo") %>% 
  sf::st_as_sf(coords = c("lon", "lat"), crs = 4326) %>% 
  ggplot()+
  tidyterra::geom_spatraster_rgb(data = bn_sat_img, maxcell = 10^7)+
  geom_sf(aes(col = habitat, size = gr_m2))+
  coord_sf(xlim = c(-78.8, -78.5), ylim = c(15.75, 15.95))+
  ggspatial::annotation_scale(location = "bl", 
                              width_hint = 0.2, 
                              style = "ticks", 
                              line_col = "white",
                              text_col = "white",
                              pad_y = unit(0.01, units =  "native"), 
                              pad_x = unit(0.01, units =  "native"))+
  ggspatial::annotation_north_arrow(location = "tr", 
                                    which_north = "true", 
                                    height = unit(0.1, "native"), width = unit(0.1, "native"),
                                    pad_x = unit(0.01, "native"), pad_y = unit(0.01, "native"),
                                    style = ggspatial::north_arrow_fancy_orienteering)+
  #paletteer::scale_color_paletteer_d("RColorBrewer::Set3")+
  scale_shape_manual(values= seq(100:110))+
  theme_light()+
  labs(col = "Habitat", size = bquote('Biomass density'(g/m^2)))+
  guides(size=guide_legend(override.aes=list(colour="grey")))+
  theme(legend.background = element_blank(),
        legend.position = "right",
        # legend.box = 'vertical',
        # legend.justification = 'left',
        # legend.box.just = 'left',
        legend.key = element_blank(),
        legend.text = element_text(colour = "black"), 
        panel.background  = element_rect(fill ="#1b2b49"),
        panel.grid = element_blank())

bn_map_spanish <- bn_map+
  PristineSeasR::scale_color_pristine_seas(palette = "alternative",
                                           labels = c("Terraza prearrecifal profunda", 
                                                "Terraza prearrecifal",
                                                "Cuenca lagunar"))+
  labs(col = "Habitat", size = bquote('Biomasa '(g/m^2)))

ggsave(bn_map_spanish,
       filename = file.path(ps_exp_path, "figures", "bajo_nuevo_fish_biomass_map_spanish.png"), 
       dpi = 300, width = 8, height = 4)
```

```{r}
(location_summary <- station_summary %>% 
   group_by(sublocation) %>% 
   summarize(across(c(richness = n_taxa, n_ind, ind_m2, gr_m2, H), mean)) %>% 
   mutate_if(is.numeric, round, 2) %>% 
   left_join(fish_data_by_transect %>% 
               group_by(sublocation) %>% 
               summarize(n_stations = n_distinct(ps_station_id),
                         n_transects = n_distinct(transect_id),
                         n_families = n_distinct(family[n_ind > 0]),
                         n_taxa = n_distinct(ps_taxon_code[n_ind > 0]))) %>% 
   select(sublocation, n_stations, n_transects, n_taxa, n_families, everything()) %>% 
   ungroup())

(habitat_summary <- station_summary %>% 
    group_by(sublocation, habitat) %>% 
    summarize(across(c(richness = n_taxa, n_ind, ind_m2, gr_m2, H), mean)) %>% 
    mutate_if(is.numeric, round, 2) %>% 
    left_join(fish_data_by_transect %>% 
                group_by(sublocation, habitat) %>% 
                summarize(n_stations = n_distinct(ps_station_id),
                          n_transects = n_distinct(transect_id),
                          n_families = n_distinct(family[n_ind > 0]),
                          n_taxa = n_distinct(ps_taxon_code[n_ind > 0]))) %>% 
    select(sublocation, n_stations, n_transects, n_taxa, n_families, everything()) %>% 
    ungroup() %>% 
    arrange(desc(richness)))
```

```{r}
fish_data_by_transect %>% 
  group_by(order, family) %>% 
  summarise(n_taxa = n_distinct(ps_taxon_code[n_ind > 0])) %>% 
  ungroup() %>% 
  arrange(desc(n_taxa))
```

Overall, we registered `r location_summary$n_taxa[location_summary$sublocation == "Bajo Nuevo"]` and `r location_summary$n_taxa[location_summary$sublocation == "Serranilla"]` distinct taxa from `r location_summary$n_families[location_summary$sublocation == "Bajo Nuevo"]` and `r location_summary$n_families[location_summary$sublocation == "Serranilla"]` families in Bajo Nuevo and Serranilla, respectively.  Families with the most number of taxa were Serranidae (n = 13), Scaridae (n = 11), Pomacentridae (n = 11), and Labridae (n = 10). On average, species richness was slightly higher in Bajo Nuevo (21 taxa per transect) than in Serranilla (18 taxa per transect). This can be partially attributed to the seagrass habitat in Serranilla which had a significantly lower number of species (1.67 species) relative to all other sites (R2 = 0.20; p = 0.001). Differences in species richness between the other habitats were not statistically significant but lagoon patch reefs in Bajo Nuevo had the highest richness (22.80 species), followed by fringing reefs in Serranilla (21.67 species). On average, species richness was higher in shallower transect (20.7 species) than in deeper ones (19.6 species) and even though differences were statistically significant (F = 4.6171; p = 0.001), depth explains little of the variance across transects (R2 = 0.02045). 

```{r, include = T, echo = F}
habitat_summary %>% 
  select(sublocation, habitat, n_stations, n_transects, n_taxa, richness, ind_m2, gr_m2) %>% 
  arrange(sublocation, desc(n_stations)) %>% 
  flextable::flextable() %>% 
  flextable::set_header_labels(values = list(sublocation = "Location", 
                                             habitat = "Habitat", 
                                             n_stations = "# stations",
                                             n_transects = "# transects",
                                             n_taxa = "# taxa",
                                             richness = "Species richness",
                                             ind_m2 = "Abundance",
                                             gr_m2 = "Biomass density")) %>% 
  flextable::merge_v(j = "sublocation") %>% 
  flextable::autofit(part = "all") %>%
  flextable::align(align = "center", part = "all")
```

```{r}
transect_summary %>% 
  group_by(depth_strata) %>% 
  summarize(across(c(n_taxa, n_ind, ind_m2, gr_m2, H), mean)) 
```

```{r}
transect_summary %>% 
  select(ind_m2, gr_m2, H, evenness, n_taxa) %>% 
  rstatix::shapiro_test(ind_m2, gr_m2, H, evenness, n_taxa)

adonis2(transect_summary %>% 
          select(n_taxa) ~  habitat + depth_strata + diver + sublocation, 
        data = transect_summary,
        strata = transect_summary$ps_station_id,
        by = "margin",
        method = "euclidean")
```

```{r}
var_labs <- c("A", "B", "C")

names(var_labs) <- c("richness", "gr_m2", "ind_m2")

univariate_plots_by_habitat <- transect_summary %>% 
  pivot_longer(c(richness, ind_m2, gr_m2), names_to = "variable") %>% 
  mutate(variable = factor(variable,      # Reordering group factor levels
                           levels = c("richness", "ind_m2", "gr_m2"))) %>% 
  ggplot()+
  geom_boxplot(aes(x = sublocation, y = value, fill = habitat))+
  facet_wrap("variable", scales = "free", nrow = 3, 
             labeller = labeller(variable = var_labs))+
  bbplot::bbc_style()+
  #paletteer::scale_fill_paletteer_d("ggsci::default_jama")+
  PristineSeasR::scale_fill_pristine_seas(palette = "alternative")+
  theme(legend.position = "bottom", plot.title.position = "plot") +
  labs(fill = "", x = "", title = "")+
  guides(fill=guide_legend(nrow=2,byrow=TRUE))

ggsave(univariate_plots_by_habitat, 
       filename = file.path(ps_exp_path, "figures", "uvc_caribe_fish_univariate_boxplots_by_habitat.png"), 
       dpi = 300, width = 11, height = 10)

univariate_plots_by_habitat_SP <- univariate_plots_by_habitat +
  PristineSeasR::scale_fill_pristine_seas(palette = "alternative",
                                          labels = c("Terraza prearrecifal profunda", 
                                               "Terraza prearrecifal",
                                               "Arrecife periferico", "Cuenca lagunar", 
                                               "Pastos marinos"))
  paletteer::scale_fill_paletteer_d("ggsci::default_jama", 
                                    ) 
univariate_plots_by_habitat_SP %>% 
  ggsave(filename = file.path(ps_exp_path, "figures", "uvc_caribe_fish_univariate_boxplots_by_habitat-spanish.png"), 
         dpi = 300, width = 12, height = 10)

univariate_plots_by_depth <- transect_summary %>% 
  pivot_longer(c(richness, ind_m2, gr_m2), names_to = "variable") %>% 
  ggplot()+
  geom_boxplot(aes(x = sublocation, y = value, fill = depth_strata))+
  facet_wrap("variable", scales = "free", nrow = 3, 
             labeller = labeller(variable = var_labs))+
  bbplot::bbc_style()+
  paletteer::scale_fill_paletteer_d("ggsci::default_jama")+
  theme(legend.position = "bottom", plot.title.position = "plot") +
  labs(fill = "", x = "", title = "Univariate statistics by habitat")
  
ggsave(univariate_plots_by_depth, 
       filename = file.path(ps_exp_path, "figures", "uvc_caribe_fish_univariate_boxplots_by_depth.png"), 
       dpi = 300, width = 11, height = 10)
```

```{r }
(caribe_fish_diversity_treemap <- caribe_fish_data %>% 
  group_by(sublocation, order, family) %>% 
  summarise(weight = n_distinct(ps_taxon_code))%>% 
  ggplot(aes(area = weight, 
             label = paste(family, weight, sep = "\n"),
             fill = order,
             subgroup = order)) +
  treemapify::geom_treemap(show.legend = T)+
  treemapify::geom_treemap_text(colour = "black", place = "middle", reflow = T, min.size = 3)+
  labs(fill = "",
       title = "")+
  facet_wrap("sublocation") +
  scale_fill_manual(values = colorRampPalette(RColorBrewer::brewer.pal(12, "Set3"))(24))+
  bbplot::bbc_style()+
  guides(fill = guide_legend(nrow=5,byrow=TRUE))+
  theme(legend.position = "bottom"))

ggsave(caribe_fish_diversity_treemap, 
       filename = file.path(ps_exp_path, "figures", "uvc_caribe_fish_diversity_treemap.png"), 
       dpi = 300, width = 12, height = 10)

caribe_fish_diversity_treemap_SP <- caribe_fish_diversity_treemap +
  labs(fill = "",
       title = "Diversidad de especies en censos visuales de peces")

ggsave(caribe_fish_diversity_treemap_SP, 
       filename = file.path(ps_exp_path, "figures", "uvc_caribe_fish_diversity_treemap_spanish.png"), 
       dpi = 300, width = 12, height = 10)
```

## Acc curves

```{r}
library(iNEXT)

acc_data_by_atoll <- fish_data_by_transect %>%
  mutate(presence = as.integer(if_else(n_ind == 0, 0, 1))) %>% 
  select(sublocation, transect_id, taxon_sci_name, presence) %>% 
  pivot_wider(names_from = transect_id, 
              values_from = presence, values_fill = 0) %>% 
  group_by(sublocation) %>% 
  group_split(keep = FALSE)  %>% 
  set_names(c("Bajo Nuevo", "Serranilla")) %>% 
  map(.f = ~column_to_rownames(.x, "taxon_sci_name")) %>% 
  map(.f = ~select_if(.x, ~ !is.numeric(.) || sum(.) != 0))%>% 
  map(.f = ~filter(.x, rowSums(.x)>0))

ac_curve_by_atoll <- iNEXT(acc_data_by_atoll, q = 0,  datatype = "incidence_raw",endpoint = 150)

accum_curve_by_atoll <- ggiNEXT(ac_curve_by_atoll)+
  labs(x = "", y = "")+
  theme(legend.position = "right")+
  guides(linetype = "none")+
  scale_color_manual(values = c("#4C2E05", "#5BC0BE"))+
  scale_fill_manual(values = c("#4C2E05", "#5BC0BE"))
```

```{r}
acc_data_by_habitat_serr <- fish_data_by_transect %>%
  mutate(presence = as.integer(if_else(n_ind == 0, 0, 1))) %>% 
  filter(sublocation == "Serranilla") %>% 
  select(habitat, transect_id, taxon_sci_name, presence) %>% 
  pivot_wider(names_from = transect_id, 
              values_from = presence, values_fill = 0) %>% 
  group_by(habitat) %>% 
  group_split(keep = F)  %>% 
  set_names(c("Terraza prearrecifal", "Arrecife periferico", "Cuenca lagunar", "Pastos marinos")) %>% 
  map(.f = ~column_to_rownames(.x, "taxon_sci_name")) %>% 
  map(.f = ~select_if(.x, ~ !is.numeric(.) || sum(.) != 0))%>% 
  map(.f = ~filter(.x, rowSums(.x)>0))

ac_curve_by_hab_serra <- iNEXT(acc_data_by_habitat_serr[1:3], q = 0, 
                               datatype = "incidence_raw", endpoint =100, se = T)

accum_curve_serra <- ggiNEXT(ac_curve_by_hab_serra)+
  labs(x = "", y = "Diversidad de especies")+
  theme(legend.position = "right")+
  guides(linetype = "none")+
  PristineSeasR::scale_color_pristine_seas(palette = "alternative",
                                           reverse = T)+
  PristineSeasR::scale_fill_pristine_seas(palette = "alternative",
                                          reverse = T)+
  ggtitle('Serranilla')
```

```{r}
acc_data_by_habitat_bn <- fish_data_by_transect %>%
  mutate(presence = as.integer(if_else(n_ind == 0, 0, 1))) %>% 
  filter(sublocation == "Bajo Nuevo") %>% 
  select(habitat, transect_id, taxon_sci_name, presence) %>% 
  pivot_wider(names_from = transect_id, 
              values_from = presence, values_fill = 0) %>% 
  group_by(habitat) %>% 
  group_split(keep = F)  %>% 
  set_names(c("Terraza prearrecifal profunda", "Terraza prearrecifal", "Cuenca lagunar")) %>% 
  map(.f = ~column_to_rownames(.x, "taxon_sci_name")) %>% 
  map(.f = ~select_if(.x, ~ !is.numeric(.) || sum(.) != 0)) %>% 
  map(.f = ~filter(.x, rowSums(.x)>0))

ac_curve_by_hab_bn <- iNEXT(acc_data_by_habitat_bn, q = 0, 
                               datatype = "incidence_raw", endpoint = 100,  se = T)

library(PristineSeasR)

accum_curve_bn <- ggiNEXT(ac_curve_by_hab_bn)+
  labs(x = "Número of transectos", y = "")+
  theme(legend.position = "right")+
  guides(linetype = "none")+
  PristineSeasR::scale_color_pristine_seas(palette = "alternative")+
  PristineSeasR::scale_fill_pristine_seas(palette = "alternative") +
  ggtitle('Bajo Nuevo')
```

```{r}
library(patchwork)

acc_curves <- accum_curve_by_atoll/accum_curve_serra/accum_curve_bn & 
  theme_minimal() + 
  theme(axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"))

ggsave(acc_curves, 
       filename = file.path(ps_exp_path, "figures", "uvc_caribe_fish_acc_curves.png"), 
       dpi = 300, width = 8, height = 10)
```

```{r}
summary_by_taxa <- fish_data_by_station %>% 
  group_by(sublocation, family, taxon_sci_name) %>% 
  summarize(across(c(n_ind, f_n_ind, ind_m2, f_ind_m2, gr_m2, f_gr_m2), mean)) %>% 
  left_join(fish_data_by_station %>% 
              group_by(sublocation, family, taxon_sci_name) %>% 
              summarize(f_obs = 100*n_distinct(ps_station_id[n_ind > 0])/n_distinct(ps_station_id))) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  mutate(f_obs = round(f_obs, 1)) %>% 
  arrange(desc(gr_m2)) 
```

Fish abundance was dominated by small reef fish at both sites. In Serranilla, (*Thalassoma bifasciatum*) and (*Halichoeres bivittatus*) accounted for 25% and 7% of all fish observations. In Bajo Nuevo (*Stegastes partitus*) and (*T. bifasciatum*) were observed in all surveys and accounted for 11% and 10% of observations, on average. Most reef fish species were observed at both atolls but several others were exclusively seen on transect on only one. For instance, nurse sharks (*G. cirratum*) and (*Chaetodon capistratus*) were seen in two thirds of transects in Bajo Nuevo but were absent in Serranilla. Similarly, seven species of hamlets (*Hypoplectrus sp*) were frequently observed in Bajo Nuevo but were never seen in Serranilla. Conversely, (*Halichoeres poeyi*) and (*Haemulon parra*) and (*Pomacanthus arcuatus*) were observed in 27%, 18%, and 18% of transects in Serranilla but were absent in Bajo Nuevo. Similarly to species richness, fish abundance was significantly lower in seagrass beds that in other habitats. Including only the common habitats between atolls, average fish abundance was higher in Serranilla (2.12 ind/m2) than Bajo Nuevo (2.02 ind/m2) but differences are not statistically significant. Fish abundance was lower is deeper transects than in shallower ones and while differences are statistically significant (p = 0.017), depth explains a modest amount of the variability (R2 = 0.065)

```{r}
exclusive_sp <- summary_by_taxa %>% 
  select(sublocation, family, taxon_sci_name, f_obs) %>% 
  pivot_wider(names_from = c(sublocation), values_from = f_obs) %>% 
  filter(Serranilla == 0 | `Bajo Nuevo` == 0) %>% 
  arrange(Serranilla)

transect_summary %>% 
  filter(!habitat %in% c("deep platform", "seagrass beds")) %>% 
  group_by(sublocation) %>% 
  summarize(across(c(richness = n_taxa, n_ind, ind_m2, gr_m2, H), mean)) 

adonis2(transect_summary %>% 
          filter(!habitat %in% c("deep platform", "seagrass beds")) %>% 
          select(ind_m2) ~  habitat + depth_strata + diver + sublocation, 
        data = transect_summary %>% 
            filter(!habitat %in% c("deep platform", "seagrass beds")) ,
        strata = transect_summary %>% 
            filter(!habitat %in% c("deep platform", "seagrass beds")) %>% 
          pull(ps_station_id),
        by = "margin",
        method = "euclidean")
```

```{r}
adonis2(station_summary %>% 
          filter(!habitat %in% c("deep platform", "seagrass beds")) %>% 
          select(ind_m2) ~  habitat + sublocation, 
        data = station_summary %>% 
            filter(!habitat %in% c("deep platform", "seagrass beds")),
        by = "margin",
        method = "euclidean")

pairwiseAdonis::pairwise.adonis2(station_summary %>% 
                                   filter(!habitat %in% c("deep platform", "seagrass beds")) %>% 
                                   select(ind_m2) ~  habitat + sublocation, 
                                 data = station_summary %>% 
                                   filter(!habitat %in% c("deep platform", "seagrass beds")),
                                 by = "margin",
                                 method = "euclidean")
```

In Serranilla, the highest average biomass density (gr/m2) was attributed to (*Haemulon carbonarium*), (*Melichthys niger*), and (*Balistes vetula*) with values of 11.1, 8.16, and 5.32 gr/m2, respectively. These species were also very common and were observed in 82%, 91%, and 91% of all transects. In Bajo Nuevo, biomass density was dominated by nurse sharks (*Ginglymostoma cirratum*) with an average of 37.2 gr/m2. Nurse sharks were observed in two thirds of all stations and on average accounted for 28.9% of biomass in a transect. After nurse sharks, (*M. niger*) and (*Sphyraena barracuda*) contributed most in terms of biomass with average densities of 10.7 and 3.7 gr/m2, respectively. Overall, average fish biomass was slightly higher in Bajo Nuevo (91.73 gr/m2) than in Serranilla (81.78 gr/m2). However, if we omit nurse sharks from the analysis and limit the comparison to shared habitats, fish biomass density is considerably higher in Serranilla (89.43) than in Bajo Nuevo (50.72), although not statistically significant. Importantly, the biomass density of predatory fish and commercially important species (i.e., *Lutjanus* sp., *"Mycteroperca* sp., *Caranx* sp., *Seriola* sp.,) was very low at both Serranilla (2.12 gr/m2) and Bajo Nuevo (0.62 gr/m2). 

```{r}
fish_data_by_transect %>% 
  filter(family != "Ginglymostomatidae" & taxon_sci_name != "Decapterus macarellus") %>% 
    filter(!habitat %in% c("deep platform", "seagrass beds")) %>% 
  group_by(sublocation, ps_station_id, transect_id, habitat, depth_strata, depth_m, diver) %>% 
  summarize(n_taxa = n_distinct(ps_taxon_code[n_ind > 0]),
            n_ind = sum(n_ind),
            ind_m2 = sum(ind_m2),
            gr_m2 = sum(gr_m2)) %>% 
  ungroup() %>% 
  group_by(sublocation, habitat, ps_station_id) %>% 
  summarize(across(c(n_taxa, n_ind, ind_m2, gr_m2), mean)) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  ungroup() %>% 
  group_by(sublocation) %>% 
  summarize(across(c(n_taxa, n_ind, ind_m2, gr_m2), mean)) 
```

```{r}
adonis2(transect_summary %>% 
          filter(!habitat %in% c("deep platform", "seagrass beds")) %>% 
          select(gr_m2) ~  habitat + depth_strata + diver + sublocation, 
        data = transect_summary %>% 
            filter(!habitat %in% c("deep platform", "seagrass beds")) ,
        strata = transect_summary %>% 
            filter(!habitat %in% c("deep platform", "seagrass beds")) %>% 
          pull(ps_station_id),
        by = "margin",
        method = "euclidean")

adonis2(station_summary %>% 
          filter(!habitat %in% c("deep platform", "seagrass beds")) %>% 
          select(gr_m2) ~  habitat + sublocation, 
        data = station_summary %>% 
            filter(!habitat %in% c("deep platform", "seagrass beds")),
        by = "margin",
        method = "euclidean")

adonis2(transect_summary %>% 
          filter(!habitat %in% c("seagrass beds", "deep platform")) %>% 
          select(gr_m2, ind_m2, n_taxa) ~ sublocation + habitat + depth_m + diver, 
        data = transect_summary %>% 
          filter(!habitat %in% c("seagrass beds", "deep platform")),
        strata = transect_summary$ps_station_id[!transect_summary$habitat %in% 
                                                  c("seagrass beds", "deep platform")],
        by = "margin")
```

```{r, include = T, echo = F}
summary_by_taxa %>% 
  select(sublocation, family, taxon_sci_name, n_ind, gr_m2, f_obs) %>% 
  group_by(sublocation) %>% 
  slice_max(gr_m2, n = 10) %>% 
  ungroup() %>% 
  pivot_longer(-c(sublocation, taxon_sci_name, family)) %>% 
  pivot_wider(names_from = c(sublocation, name), values_from = value) %>% 
  arrange(family, taxon_sci_name) %>% 
  flextable::flextable() %>% 
  flextable::set_header_labels(values = list(family = "Family", 
                                             taxon_sci_name = "Taxa", 
                                             `Bajo Nuevo_n_ind` = "Abundance",
                                             `Bajo Nuevo_gr_m2` = "Biomass density",
                                             `Bajo Nuevo_f_obs` = "% Ocurrence",
                                             `Serranilla_n_ind` = "Abundance",
                                             `Serranilla_gr_m2` = "Biomass density",
                                             `Serranilla_f_obs` = "% Ocurrence")) %>% 
  flextable::add_header_row(values = c("", "", "Bajo Nuevo", "Serranilla"), 
                            colwidths = c(1,1,3,3)) %>% 
  flextable::merge_v(j = "family") %>% 
  flextable::autofit(part = "all") %>%
  flextable::align(align = "center", part = "all")
```

```{r}
caribe_fish_biomass_treemap <- fish_data_by_transect %>% 
  #filter(family != "Ginglymostomatidae") %>% 
  group_by(sublocation, ps_station_id, ps_taxon_code, order, family) %>% 
  summarise(across(c(ind_m2, gr_m2), mean)) %>%
  ungroup() %>% 
  group_by(sublocation, order, family) %>% 
  summarise(across(c(ind_m2, gr_m2), sum)) %>% 
  ungroup() %>% 
  mutate(f_ind_m2 = round(100*ind_m2/sum(ind_m2),1),
         f_gr_m2 = round(100*gr_m2/sum(gr_m2), 1)) %>% 
  ggplot(aes(area = gr_m2, 
             label = paste(family, paste0(f_gr_m2, "%"), sep = "\n"),
             fill = order,
             subgroup = order)) +
  treemapify::geom_treemap(show.legend = T)+
  treemapify::geom_treemap_text(colour = "white", place = "middle", reflow = T, min.size = 3)+
  labs(fill = "",
       title = "Taxa biomass (gr m2) in underwater fish surveys")+
  facet_wrap("sublocation", nrow = 1) +
  scale_fill_manual(values = colorRampPalette(RColorBrewer::brewer.pal(12, "Set3"))(24))+
  bbplot::bbc_style()+
  guides(fill = guide_legend(nrow=5,byrow=TRUE))+
  theme(legend.position = "bottom")

ggsave(caribe_fish_biomass_treemap, 
       filename = file.path(ps_exp_path, "figures", "uvc_caribe_fish_biomass_treemap.png"),
       dpi = 300, width = 12, height = 10)
```

```{r}
fish_data_by_transect %>% 
  filter(str_detect(taxon_sci_name, c("Caranx", "Mycteroperca", "Seriola", "Lutjanus"))) %>% 
  group_by(sublocation, ps_station_id, transect_id, habitat, depth_strata, depth_m, diver) %>% 
  summarize(n_taxa = n_distinct(ps_taxon_code[n_ind > 0]),
            n_ind = sum(n_ind),
            ind_m2 = sum(ind_m2),
            gr_m2 = sum(gr_m2)) %>% 
  ungroup() %>% 
  group_by(sublocation, habitat, ps_station_id) %>% 
  summarize(across(c(n_taxa, n_ind, ind_m2, gr_m2), mean)) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  ungroup() %>% 
  group_by(sublocation) %>% 
  summarize(across(c(n_taxa, n_ind, ind_m2, gr_m2), mean)) 
```

## Community composition

Fish assemblages were significantly different between atolls (R2 = 0.16; F = 4.75; p= 0.001). Stations in Serranilla show a higher concordance in the ordination space relative to Bajo Nuevo which is expected given the smaller sampling footprint in the latter atoll. Species that strongly determine the ordination of stations include the nurse sharks (*G. cirratum*) and hamlet (*H. nigricans*) which were abundant in Bajo Nuevo but absent in Serranilla, and (*Balistes vetula*) whose biomass density was twice as high in Serranilla compared to Bajo Nuevo. 

```{r}
gr_m2_data_wide <- fish_data_by_station %>% 
  filter(!habitat %in% c("deep platform", "seagrass beds")) %>% 
  select(ps_station_id, taxon_sci_name, gr_m2) %>% 
  pivot_wider(names_from = taxon_sci_name, values_from = gr_m2, values_fill = 0) %>% 
  column_to_rownames("ps_station_id") 

gr_m2_dist <- vegan::vegdist(as.matrix(gr_m2_data_wide), 
                      method = "bray") 

gr_m2_pcoa_c <- dbrda(gr_m2_dist ~ habitat + sublocation, 
                      data = station_summary %>% 
                        filter(!habitat %in% c("deep platform", "seagrass beds")),
                      dist = "bray", 
                      add = "lingoes")

gr_m2_pcoa_c_env_fit <- envfit(gr_m2_pcoa_c, 
                               station_summary %>% 
                                 filter(!habitat %in% c("deep platform", "seagrass beds")) %>% 
                                 select(habitat,  sublocation))

gr_m2_pcoa_c_spp_fit <- envfit(gr_m2_pcoa_c, 
                               gr_m2_data_wide)

gr_m2_pcoa_c_spp_scores <- gr_m2_pcoa_c_spp_fit %>% 
  scores("vectors") %>% 
  as_tibble(rownames = "ps_taxon_code") %>% 
  mutate(r = gr_m2_pcoa_c_spp_fit$vectors$r,
         p = gr_m2_pcoa_c_spp_fit$vectors$pvals) %>% 
  filter(p < 0.05, r > 0.1)

gr_m2_pcoa_c_axis <- BiodiversityR::axis.long(gr_m2_pcoa_c, choices = c(1, 2))

gr_m2_pcoa_c_site_scores <- scores(gr_m2_pcoa_c)$sites %>% 
  as_tibble(rownames = "ps_station_id") %>% 
  inner_join(station_summary)
```

```{r}
community_pcoa_plots <- map(c("habitat", "sublocation"), 
    ~ ggplot() +
      geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
      geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
      scale_x_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      geom_point(data = gr_m2_pcoa_c_site_scores,
                 aes(x = dbRDA1, y = dbRDA2,, col = .data[[.x]])) +
      labs(x = gr_m2_pcoa_c_axis$label[1], y = gr_m2_pcoa_c_axis$label[2])+
      geom_segment(data = gr_m2_pcoa_c_spp_scores %>% 
                     filter(p < 0.05, r > 0.5), 
                   aes(x = 0, y = 0, xend = dbRDA1*3, yend = dbRDA2*3, alpha = r), 
                   colour = "black", 
                   size = 0.2, 
                   arrow = arrow(length = unit(0.01, "npc"), 
                                 type = 'open', ends = "last"))+
      ggrepel::geom_text_repel(data = gr_m2_pcoa_c_spp_scores %>% 
                                 filter(p < 0.05, r > 0.5), 
                               aes(x=dbRDA1*3, y=dbRDA2*3, label = ps_taxon_code, alpha = r),
                               colour="black",
                               show.legend = F) +
      coord_fixed(ratio=1)+
      labs(color = "", alpha = bquote('R'^'2'))+
      ggforce::geom_mark_ellipse(data = gr_m2_pcoa_c_site_scores, 
                                 aes(x=dbRDA1, y = dbRDA2, colour = .data[[.x]], 
                                     fill=after_scale(alpha(colour, 0.01))), 
                                 expand=0, size=0.2, show.legend=FALSE)+
      theme(panel.background = element_blank(),
            panel.border = element_blank(),
            panel.grid = element_blank(),
            axis.line = element_line("gray25"),
            text = element_text(size = 12, family = "Arial"),
            axis.text = element_text(size = 10, colour = "gray25"),
            axis.title = element_text(size = 12, colour = "gray25"),
            legend.key = element_blank())+
      PristineSeasR::scale_color_pristine_seas(palette = "alternative")+
      scale_alpha(range=c(0.6,1)))

pcoa_by_hab <- community_pcoa_plots[[1]] +
  labs(color = "Habitat") + 
  PristineSeasR::scale_color_pristine_seas(palette = "alternative",
                                           labels = c("Terraza prearrecifal", 
                                                "Arrecife periferico","Cuenca lagunar"))
 
pcoa_by_location <- community_pcoa_plots[[2]]+ 
  scale_color_manual(values = c("#4C2E05", "#5BC0BE"))
  
library(patchwork)

pcoa_plots <- pcoa_by_location/pcoa_by_hab 
  
ggsave(pcoa_plots, filename = file.path(ps_exp_path, 
                                        "figures", "uvc_caribe_fish_pcoa_plots.png"), 
       dpi = 300, width = 8, height = 10)

# ggsave(pcoa_by_hab, 
#        filename = file.path(ps_exp_path, "figures", "uvc_caribe_fish_biomas_pcoa_by_habitat_spanish.png"), 
#        dpi = 300, width = 8, height = 6)
# 
# 
# pcoa_by_location <- community_pcoa_plots[[2]] 
# 
#   ggsave(filename = file.path(ps_exp_path, "figures", "uvc_caribe_fish_biomas_pcoa_by_sublocation.png"), 
#        dpi = 300, width = 8, height = 6)
```

```{r}
anova(gr_m2_pcoa_c) # overall test of the significant of the analysis
anova(gr_m2_pcoa_c, by = "axis", perm.max = 500) # test axes for significance
anova(gr_m2_pcoa_c, by = "terms", perm.max = 1000) # test for sign. environ. variables
```

```{r}
n_ind_adonis <- adonis2(gr_m2_dist ~  habitat  + sublocation,
        data = station_summary[!station_summary$habitat %in% c("deep platform", "seagrass beds"), ],
        #strata = transect_summary[transect_summary$n_ind > 10, ]$ps_station_id,
        by = "margin")

n_ind_adonis
```


### PCOA

```{r, eval = F}
min_ok_n <- 10

max_n_samp <- min(transect_summary$n_ind[transect_summary$n_ind > min_ok_n])

n_ind_data_wide_trimmed <- n_ind_data_wide[transect_summary$transect_id[transect_summary$n_ind > min_ok_n],]

cols_to_rm <- colSums(n_ind_data_wide[row.names(n_ind_data_wide_trimmed),]) == 0
       
n_ind_data_wide_trimmed <- n_ind_data_wide_trimmed[,!cols_to_rm] 

set.seed(19900623)

n_ind_dist <- avgdist(as.matrix(n_ind_data_wide_trimmed), 
                      dmethod = "bray", 
                      sample = max_n_samp) 
```

#### Unconstrained

```{r, eval = F}
fish_pcoa <- dbrda(n_ind_dist ~ 1, 
                   data = transect_summary[transect_summary$n_ind > min_ok_n, ],
                   add = "lingoes")

fish_pcoa_env_fit <- envfit(fish_pcoa, 
                            transect_summary %>% 
                              filter(n_ind > min_ok_n) %>% 
                              select(habitat, depth_strata, diver, sublocation))

fish_pcoa_spp_fit <- envfit(fish_pcoa, 
                            n_ind_data_wide_trimmed)

fish_pcoa_spp_scores <- fish_pcoa_spp_fit %>% 
  scores("vectors") %>% 
  as_tibble(rownames = "ps_taxon_code") %>% 
  mutate(r = fish_pcoa_spp_fit$vectors$r,
         p = fish_pcoa_spp_fit$vectors$pvals) %>% 
  filter(p < 0.05, r > 0.1)

fish_pcoa_axis <- BiodiversityR::axis.long(fish_pcoa, choices = c(1, 2))

fish_pcoa_site_scores <- scores(fish_pcoa) %>% 
  as_tibble(rownames = "transect_id") %>% 
  inner_join(transect_summary)
```

```{r, eval = F}
map(c("habitat", "sublocation", "diver", "depth_strata"), 
    ~ ggplot() +
      geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
      geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
      scale_x_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      geom_point(data = fish_pcoa_site_scores,
                 aes(x = MDS1, y = MDS2,, col = .data[[.x]])) +
      labs(x = fish_pcoa_axis$label[1], y = fish_pcoa_axis$label[2])+
      geom_segment(data = fish_pcoa_spp_scores %>% 
                     filter(p < 0.05, r > 0.2), 
                   aes(x = 0, y = 0, xend = MDS1*3, yend = MDS2*3, alpha = r), 
                   colour = "black", 
                   size = 0.2, 
                   arrow = arrow(length = unit(0.01, "npc"), 
                                 type = 'open', ends = "last"))+
      ggrepel::geom_text_repel(data = fish_pcoa_spp_scores %>% 
                                 filter(p < 0.05, r > 0.2), 
                               aes(x=MDS1*3, y=MDS2*3, label = ps_taxon_code, alpha = r),
                               colour="black",
                               show.legend = F) +
      coord_fixed(ratio=1)+
      ggforce::geom_mark_ellipse(data = fish_pcoa_site_scores, 
                                 aes(x=MDS1, y = MDS2, colour = .data[[.x]], 
                                     fill=after_scale(alpha(colour, 0.01))), 
                                 expand=0, size=0.2, show.legend=FALSE)+
      theme(panel.background = element_blank(),
            panel.border = element_blank(),
            panel.grid = element_blank(),
            axis.line = element_line("gray25"),
            text = element_text(size = 12, family = "Arial"),
            axis.text = element_text(size = 10, colour = "gray25"),
            axis.title = element_text(size = 12, colour = "gray25"),
            legend.key = element_blank())+
      scale_alpha(range=c(0.6,1))
)
```

```{r, eval = F}
fish_pcoa_env_fit
```

#### Constrained

```{r, eval = F}
fish_pcoa_c <- dbrda(n_ind_dist ~ diver + habitat + sublocation + depth_strata, 
                    data = transect_summary[transect_summary$n_ind > min_ok_n, ],
                    dist = "bray", 
                    add = "lingoes")

fish_pcoa_c_env_fit <- envfit(fish_pcoa_c, 
                            transect_summary %>% 
                              filter(n_ind > min_ok_n) %>% 
                              select(habitat, depth_strata, diver, sublocation))

fish_pcoa_c_spp_fit <- envfit(fish_pcoa_c, 
                            n_ind_data_wide_trimmed)

fish_pcoa_c_spp_scores <- fish_pcoa_c_spp_fit %>% 
  scores("vectors") %>% 
  as_tibble(rownames = "ps_taxon_code") %>% 
  mutate(r = fish_pcoa_c_spp_fit$vectors$r,
         p = fish_pcoa_c_spp_fit$vectors$pvals) %>% 
  filter(p < 0.05, r > 0.1)

fish_pcoa_c_axis <- BiodiversityR::axis.long(fish_pcoa_c, choices = c(1, 2))

fish_pcoa_c_site_scores <- scores(fish_pcoa_c)$sites %>% 
  as_tibble(rownames = "transect_id") %>% 
  inner_join(transect_summary)

#summary(fish_pcoa_c)
```

```{r, eval = F}
map(c("habitat", "sublocation", "diver", "depth_strata"), 
    ~ ggplot() +
      geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
      geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
      scale_x_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      geom_point(data = fish_pcoa_c_site_scores,
                 aes(x = dbRDA1, y = dbRDA2,, col = .data[[.x]])) +
      labs(x = fish_pcoa_c_axis$label[1], y = fish_pcoa_c_axis$label[2])+
      geom_segment(data = fish_pcoa_c_spp_scores %>% 
                     filter(p < 0.05, r > 0.2), 
                   aes(x = 0, y = 0, xend = dbRDA1*5, yend = dbRDA2*5, alpha = r), 
                   colour = "black", 
                   size = 0.2, 
                   arrow = arrow(length = unit(0.01, "npc"), 
                                 type = 'open', ends = "last"))+
      ggrepel::geom_text_repel(data = fish_pcoa_c_spp_scores %>% 
                                 filter(p < 0.05, r > 0.2), 
                               aes(x=dbRDA1*5, y=dbRDA2*5, label = ps_taxon_code, alpha = r),
                               colour="black",
                               show.legend = F) +
      coord_fixed(ratio=1)+
      ggforce::geom_mark_ellipse(data = fish_pcoa_c_site_scores, 
                                 aes(x=dbRDA1, y = dbRDA2, colour = .data[[.x]], 
                                     fill=after_scale(alpha(colour, 0.01))), 
                                 expand=0, size=0.2, show.legend=FALSE)+
      theme(panel.background = element_blank(),
            panel.border = element_blank(),
            panel.grid = element_blank(),
            axis.line = element_line("gray25"),
            text = element_text(size = 12, family = "Arial"),
            axis.text = element_text(size = 10, colour = "gray25"),
            axis.title = element_text(size = 12, colour = "gray25"),
            legend.key = element_blank())+
      scale_alpha(range=c(0.6,1))
)
```

```{r, eval = F}
anova(fish_pcoa_c) # overall test of the significant of the analysis
anova(fish_pcoa_c, by = "axis", perm.max = 500) # test axes for significance
anova(fish_pcoa_c, by = "terms", perm.max = 10000) # test for sign. environ. variables
```

### NMDS

```{r, eval = F}
n_ind_MDS <- metaMDS(n_ind_dist, k = 2, trymax = 1000)

n_ind_MDS$stress

stressplot(n_ind_MDS)

n_ind_MDS_env_fit <- envfit(n_ind_MDS, 
                            transect_summary %>% 
                              filter(n_ind > min_ok_n) %>% 
                              select(habitat, depth_strata, diver, sublocation))

n_ind_MDS_spp_fit <- envfit(n_ind_MDS, 
                            n_ind_data_wide_trimmed)

n_ind_MDS_spp_scores <- n_ind_MDS_spp_fit %>% 
  scores("vectors") %>% 
  as_tibble(rownames = "ps_taxon_code") %>% 
  mutate(r = n_ind_MDS_spp_fit$vectors$r,
         p = n_ind_MDS_spp_fit$vectors$pvals) %>% 
  filter(p < 0.05, r > 0.1)

n_ind_MDS_axis <- BiodiversityR::axis.long(n_ind_MDS, choices = c(1, 2))

n_ind_MDS_site_scores <- scores(n_ind_MDS, display = "sites") %>% 
  as_tibble(rownames = "transect_id") %>% 
  inner_join(transect_summary)
```

```{r, eval = F}
map(c("habitat", "sublocation", "diver", "depth_strata"), 
    ~ ggplot() +
      geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
      geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
      scale_x_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      geom_point(data = n_ind_MDS_site_scores,
                 aes(x = NMDS1, y = NMDS2,, col = .data[[.x]])) +
      labs(x = n_ind_MDS_axis$label[1], y = n_ind_MDS_axis$label[2])+
      geom_segment(data = n_ind_MDS_spp_scores %>% 
                     filter(p < 0.05, r > 0.2), 
                   aes(x = 0, y = 0, xend = NMDS1*1, yend = NMDS2*1, alpha = r), 
                   colour = "black", 
                   size = 0.2, 
                   arrow = arrow(length = unit(0.01, "npc"), 
                                 type = 'open', ends = "last"))+
      ggrepel::geom_text_repel(data = n_ind_MDS_spp_scores %>% 
                                 filter(p < 0.05, r > 0.2), 
                               aes(x=NMDS1*1, y=NMDS2*1, label = ps_taxon_code, alpha = r),
                               colour="black",
                               show.legend = F) +
      coord_fixed(ratio=1)+
      ggforce::geom_mark_ellipse(data = n_ind_MDS_site_scores, 
                                 aes(x=NMDS1, y = NMDS2, colour = .data[[.x]], 
                                     fill=after_scale(alpha(colour, 0.01))), 
                                 expand=0, size=0.2, show.legend=FALSE)+
      theme(panel.background = element_blank(),
            panel.border = element_blank(),
            panel.grid = element_blank(),
            axis.line = element_line("gray25"),
            text = element_text(size = 12, family = "Arial"),
            axis.text = element_text(size = 10, colour = "gray25"),
            axis.title = element_text(size = 12, colour = "gray25"),
            legend.key = element_blank())+
      scale_alpha(range=c(0.6,1))
)
```

```{r, eval = F}
n_ind_adonis <- adonis2(n_ind_dist ~ diver + habitat + depth_strata + sublocation,
        data = transect_summary[transect_summary$n_ind > 10, ],
        strata = transect_summary[transect_summary$n_ind > 10, ]$ps_station_id,
        by = "margin")

n_ind_adonis
```

```{r, eval = F}
sppscores(n_ind_MDS) <- n_ind_data_wide_trimmed

n_ind_MDS_spp_scores_2 <- scores(n_ind_MDS)$species %>% 
  as_tibble(rownames = "ps_taxon_code") 

map(c("habitat", "sublocation", "diver", "depth_strata"), 
    ~ ggplot() +
      geom_vline(xintercept = c(0), color = "grey70", linetype = 2) +
      geom_hline(yintercept = c(0), color = "grey70", linetype = 2) +  
      scale_x_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      scale_y_continuous(sec.axis = dup_axis(labels=NULL, name=NULL)) +
      geom_point(data = n_ind_MDS_site_scores,
                 aes(x = NMDS1, y = NMDS2,, col = .data[[.x]])) +
      labs(x = n_ind_MDS_axis$label[1], y = n_ind_MDS_axis$label[2])+
      geom_segment(data = n_ind_MDS_spp_scores_2 , 
                   aes(x = 0, y = 0, xend = NMDS1*.5, yend = NMDS2*.5), 
                   colour = "black", 
                   size = 0.2, 
                   arrow = arrow(length = unit(0.01, "npc"), 
                                 type = 'open', ends = "last"))+
      ggrepel::geom_text_repel(data = n_ind_MDS_spp_scores_2, 
                               aes(x=NMDS1*.5, y=NMDS2*.5, label = ps_taxon_code),
                               colour="black",
                               show.legend = F) +
      coord_fixed(ratio=1)+
      ggforce::geom_mark_ellipse(data = n_ind_MDS_site_scores, 
                                 aes(x=NMDS1, y = NMDS2, colour = .data[[.x]], 
                                     fill=after_scale(alpha(colour, 0.01))), 
                                 expand=0, size=0.2, show.legend=FALSE)+
      theme(panel.background = element_blank(),
            panel.border = element_blank(),
            panel.grid = element_blank(),
            axis.line = element_line("gray25"),
            text = element_text(size = 12, family = "Arial"),
            axis.text = element_text(size = 10, colour = "gray25"),
            axis.title = element_text(size = 12, colour = "gray25"),
            legend.key = element_blank())+
      scale_alpha(range=c(0.6,1))
)

```

# Additional figs

```{r, eval = F}
transect_summary %>% 
  pivot_longer(c(ind_m2, gr_m2, H, richness, evenness)) %>% 
  ggplot()+
  geom_boxplot(aes(x = sublocation, y = value, fill = diver))+
  facet_wrap("name", scales = "free")

transect_summary %>% 
  pivot_longer(c(ind_m2, gr_m2, H, richness, evenness)) %>% 
  ggplot()+
  geom_boxplot(aes(x = sublocation, y = value, fill = habitat))+
  facet_wrap("name", scales = "free")

transect_summary %>% 
  pivot_longer(c(ind_m2, gr_m2, H, richness, evenness)) %>% 
  ggplot()+
  geom_boxplot(aes(x = sublocation, y = value, fill = depth_strata))+
  facet_wrap("name", scales = "free")
```


```{r, include = T}
master_taxa <- read_csv(file.path(ps_exp_path, "data", "primary", "processed","taxa", "clean_master_taxa.csv")) %>% 
  filter(location == "Caribbean")

dscm_elasmo <- c("Heptranchias perlo",
                 "Galeus springeri",
                 "Mustelus canis",
                 "Alopias superciliosus") %>% 
  furrr::future_map_dfr(worrms::wm_records_names) %>% 
  janitor::clean_names() %>% 
  filter(!is.na(scientificname)) %>% 
  mutate(taxon_id = if_else(!is.na(valid_aphia_id),
                              paste0("WORMS:", valid_aphia_id),
                              NA_character_)) %>% 
  select(scientificname, status, taxon_valid_name = valid_name, taxon_rank = rank, taxon_id, taxon_authority = valid_authority,
         kingdom, phylum, class, order, family, genus) %>% 
  distinct() 

master_taxa %>% 
  filter(class == "Elasmobranchii", taxon_rank == "Species") %>% 
  select(class, order, family, genus,taxon_sci_name, iucn_redlist_cat, iucn_trend) %>% 
  distinct() %>% 
  filter(!is.na(iucn_redlist_cat)) %>% 
  bind_rows(dscm_elasmo %>% 
              select(class, order, family, genus, taxon_sci_name = taxon_valid_name)) %>% 
  arrange(class, order, family) %>%
  select(-class, -genus) %>% 
  flextable::flextable() %>% 
  flextable::merge_v(j = c("order", "family")) %>% 
  flextable::autofit(part = "all") %>%
  flextable::align(align = "center", part = "all")
```

```{r, include = T}
master_taxa %>% 
  filter(!is.na(iucn_redlist_cat), 
         !iucn_redlist_cat %in%c("Least Concern", "Data Deficient")) %>% 
  select(phylum, class, order, family, taxon_sci_name, iucn_redlist_cat, iucn_trend) %>% 
  distinct() %>% 
  arrange(class, order, family) %>%
  flextable::flextable() %>% 
  flextable::merge_v(j = c("phylum","class","order", "family")) %>% 
  flextable::autofit(part = "all") %>%
  flextable::align(align = "center", part = "all")
```


