---
title: "Process sub in Bajo Nuevo"
author: 
output:
  word_document:
    toc: no
---

```{r echo = FALSE, include = FALSE, warning = FALSE,prompt = FALSE, error = FALSE, strip.white = TRUE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE,
                      prompt = FALSE, error = FALSE, strip.white = TRUE)

knitr::opts_knit$set(progress = TRUE, verbose = TRUE)

options(dplyr.summarise.inform = FALSE)

library(flextable)
library(raster)
library(sf)
library(ggspatial)
library(lubridate)
library(bdc)
library(terra)
library(tidyverse)
source("_prj_vars.R")

knitr::opts_chunk$set(echo = TRUE, fig.cap = TRUE)
  set_flextable_defaults(
  font.family = "Arial", font.size = 9, 
  theme_fun = "theme_vanilla",
  big.mark="", table.layout="autofit")
  
load(file = file.path(ps_exp_path, "data", "secondary", "processed", "admin_spatial_layers.RData"))

sub_meta <- read_csv(file.path(ps_exp_path, 
                               "data", "primary", "processed", "metadata", "sub_meta.csv")) %>% 
  filter(location == "Caribbean")

sub_taxa <- read_csv(file.path(ps_exp_path, 
                               "data", "primary", "processed", "taxa", "clean_master_taxa.csv")) %>% 
  filter(method == "sub", location == "Caribbean")
```

```{r, include = T, echo = F}
bn_fish_obs <- readxl::read_xlsx(file.path(ps_exp_path, "data/primary/raw",
                                              "sub","files/bajo_nuevo_annotations.xlsx"),
                  sheet = "fish obs") %>% 
  filter(!is.na(taxa)) %>% 
  select(-notes) %>% 
  mutate(time = format(time, format = "%H:%M:%S"),
         ps_station_id = if_else(ps_station_id == "SEA_sub_01", "COL_sub_19", "COL_sub_20"))

bn_fish_obs <- bn_fish_obs %>% 
  left_join(sub_taxa %>% 
              select(taxa = taxon, taxon_sci_name, family, order)) %>% 
  filter(!is.na(family)) 
```

Weather conditions allowed us to make only two dives with the DeepSee submersible in Bajo Nuevo at maximum depths of 150 and 300 meters respectively. A gently sloping shelf quickly transitioned to a vertical wall starting around 80 meters and continuing well beyond 300 meters. The vertical wall was predominantly covered by sponges and encrusting algae with some interspersed soft corals. 

Crevices and cracks in the wall provided shelter for several fish species. We recorded a total of `r n_distinct(bn_fish_obs$taxa)` different fish taxa belonging to `r n_distinct(bn_fish_obs$family)` families. The most abundant fish species was the blackcap basslet (*Gramma melacara*), followed by (*Xanthicthys ringens*) and (*Chromis insolata*). Fish species richness and abundance decreased with depth and only five species were observed deeper than 200 meters. Notably, two of these species: the french butterflyfish (*Prognathodes guyanensis*) and the golden bass (*Lipoproma aberrans*) represent new records for the archipelago of San Andres and Old Providence. Also, the Spanish flag (*Gonioplectrus hispanus*), observed at 177 meters, and Northern red snapper (*Lutjanus campechanus*) observed at 150 meters, are new additions to the inventory of fish species of the archipelago. 

```{r}
bn_fish_obs %>% 
  filter(depth > 200) %>% 
  group_by(taxa) %>% 
  summarize(min_depth = min(depth), 
            max_depth = max(depth), 
            sum_ind = sum(n_ind)) %>% 
  arrange(desc(sum_ind)) %>% 
  flextable() %>% 
  merge_v(j = "family") %>% 
  autofit(part = "all") %>%
  flextable::align(align = "center", part = "all")
```

```{r, include = T, echo = F}
sub_bajo_nuevo_barplot <- bn_fish_obs %>% 
  filter(depth > 80) %>% 
  mutate(depth_bin = cut(depth, seq(0, 300, by = 20))) %>% 
  group_by(depth_bin) %>% 
  summarize(n_taxa = n_distinct(taxa),
            n_ind = sum(n_ind)) %>% 
  pivot_longer(!depth_bin) %>% 
  ggplot()+
  geom_col(aes(x = depth_bin, y = value, fill = name), position = "dodge")+
  labs(y = "", x = "Depth (m)")+
  paletteer::scale_fill_paletteer_d("awtools::mpalette", labels = c("# individuos", "# taxones"))+
  bbplot::bbc_style()

ggsave(plot = sub_bajo_nuevo_barplot, 
       filename = file.path(ps_exp_path, "figures", "sub_bajo_nuevo_fish_barplot.png"),
        width = 12)
```

```{r}
peces_sai <- readxl::read_excel(file.path(ps_exp_path, "data/secondary/raw/biodiversidad del mar de los 7 colores", "PECES_REV.xlsx")) 
```

```{r}
sub_taxa %>% 
  filter(taxon_rank == "Species", !taxon_sci_name %in% peces_sai$Especie)
```
```{r include = T, echo = F}
sub_meta %>% 
  select(date, ps_station_id, dive_start, bottom_time, lat = surface_start_lat, lon = surface_start_lon, max_depth, temp_at_max_depth) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  flextable() %>% 
  merge_v(j = "family") %>% 
  autofit(part = "all") %>%
  flextable::align(align = "center", part = "all")
```

