---
title: "Process seabird"
author: 
output:
  word_document:
    toc: no
---

```{r echo = FALSE, include = FALSE, warning = FALSE,prompt = FALSE, error = FALSE, strip.white = TRUE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE,
                      prompt = FALSE, error = FALSE, strip.white = TRUE)

knitr::opts_knit$set(progress = TRUE, verbose = TRUE)

options(dplyr.summarise.inform = FALSE)

library(tidyverse)
library(flextable)
library(raster)
library(sf)
library(ggspatial)
library(lubridate)
library(bdc)
library(terra)
source("_prj_vars.R")

knitr::opts_chunk$set(echo = TRUE, fig.cap = TRUE)
  set_flextable_defaults(
  font.family = "Arial", font.size = 9, 
  theme_fun = "theme_vanilla",
  big.mark="", table.layout="autofit")
  
load(file = file.path(ps_exp_path, "data", "secondary", "processed", "admin_spatial_layers.RData"))

birds_fieldsheet <- readxl::read_xlsx(file.path(ps_exp_path, 
                                          "data/primary/raw",
                                          "birds","Colombia_Seabirds_Fieldsheet_2022_05_17_lat_longs updated.xlsx"), 
                                sheet = "Transect meta") %>% 
  janitor::clean_names() %>% 
  filter(location == "Seaflower") %>% 
  mutate(location = "Caribbean")

birds_meta <- birds_fieldsheet %>% 
  transmute(location, sublocation, ps_station_id, 
            method = "Seabird transects",
            date = ymd(date),
            time_start = hms::as_hms(time_start),
            duration_mins,
            lat_start,
            long_start,
            lat_stop = as.numeric(lat_stop),
            long_stop,
            distance_km, 
            paired_pelagic, 
            string, 
            drop_pickup, 
            vessel, 
            observer = "Chris Thompson") %>% 
    mutate(location = case_when(str_detect(sublocation, "Transit") ~ "Transit",
                              str_detect(sublocation, "Bajo Nuevo") ~ "Bajo Nuevo",
                              str_detect(sublocation, "Serranilla") ~ "Serranilla",
                              TRUE ~ NA_character_)) 

birds_census <- readxl::read_xlsx(file.path(ps_exp_path, 
                            "data/primary/raw",
                            "birds","Colombia_Seabirds_Fieldsheet_2022_05_17_lat_longs updated.xlsx"), 
                  sheet = "Transect meta") %>% 
  filter(Location == "Seaflower") %>% 
  dplyr::select(-c(1:16)) %>% 
  mutate_all(as.numeric) %>% 
  mutate(ps_station_id = birds_meta$ps_station_id,
         sublocation = birds_meta$sublocation) %>% 
  pivot_longer(-c("ps_station_id", "sublocation"), names_to = "taxa", values_to = "n_ind") %>% 
  mutate(taxa = if_else(str_detect(taxa, "red-footed"), "Red footed booby", taxa)) %>% 
  group_by(sublocation, ps_station_id, taxa) %>% 
  summarize(n_ind = sum(n_ind)) %>% 
  filter(n_ind > 0, !taxa %in% c("shorebird sp", "grey mid-size petrel")) %>% 
  ungroup() %>% 
  mutate(location = case_when(str_detect(sublocation, "Transit") ~ "Transit",
                              str_detect(sublocation, "Bajo Nuevo") ~ "Bajo Nuevo",
                              str_detect(sublocation, "Serranilla") ~ "Serranilla",
                              TRUE ~ NA_character_)) %>% 
  dplyr::select(location, everything()) %>% 
  rename(common_name = taxa) %>% 
  mutate(common_name = str_to_sentence(common_name))
```

```{r, eval = F}
birds_taxa <- birds_census %>% 
  distinct(common_name) %>% 
  left_join(readxl::read_xlsx(file.path(ps_exp_path, 
                                        "data/primary/raw",
                                        "birds","Colombia_Seabirds_Fieldsheet_2022_05_17_lat_longs updated.xlsx"), 
                              sheet = "Species Seaflower") %>% 
              janitor::clean_names() %>% 
              dplyr::select(common_name = taxa, binomial) %>% 
              mutate(common_name = str_to_sentence(common_name))
              ) %>% 
  dplyr::select(common_name, sci_name = binomial)

birds_taxa$sci_name[birds_taxa$common_name == "Common noddy"] <- "Anous stolidus"
birds_taxa$sci_name[birds_taxa$common_name == "Smooth-billed ani"] <- "Crotophaga ani"
birds_taxa$sci_name[birds_taxa$common_name == "Blue-winged teal"] <- "Anas discors"

birds_taxa_gbif <- bdc_query_names_taxadb(birds_taxa$sci_name,
                                         replace_synonyms = TRUE,
                                         suggest_names = TRUE,
                                         suggestion_distance = 0.8,
                                         db = "gbif",
                                         parallel = TRUE,
                                         ncores = 2,
                                         export_accepted = FALSE) %>% 
  janitor::clean_names()

birds_taxa <- birds_taxa_gbif %>% 
  dplyr::select(original_search, scientific_name, family, order) %>% 
  left_join(birds_taxa, 
            by = c("original_search" = "sci_name")) 

write_csv(birds_taxa, "birds_taxa.csv")
```

```{r}
birds_taxa <- read_csv("birds_taxa.csv")
```


We recorded the abundance of seabird species present during the deployment and recovery of each of our pelagic BRUVS (Baited Remote Underwater Video System) sets, to develop a record of seabird species presence and abundance. We also made additional opportunistic records of seabird species presence off transect. In `r nrow(birds_meta[birds_meta$location == "Serranilla", ])` transects conducted in Serranilla, `r nrow(birds_meta[birds_meta$location == "Bajo Nuevo", ])` in Bajo Nuevo, and `r nrow(birds_meta[!birds_meta$location == "Transit", ])` while in transit, we observed `r nrow(birds_taxa)` species of birds belonging to `r n_distinct(birds_taxa$family)` families. While some species were observed in both atolls - the magnificent frigate (*Fregata magnificens*) and the peregrine falcon (*Falco peregrinus*) - several others were found in only one of the two. In Serranilla, the barn swallow (*Hirundo rustica*) and brown noddy (*Anous stolidus*), were recorded in 12% of surveys but were absent in Bajo Nuevo. Additionally, the smooth-billed ani (*Crotophaga ani*)  and the bank swallow (*Riparia riparia*) were exclusive to Serranilla's Cay. Conversely, the black-faced petrel(*Pterodroma hasitata*) was exclusive to Bajo Nuevo where it was observed in 30% of the surveys and so was the glossy ibis (*Plegadis falcinellus*); which was seen once but in a relatively large flock (19 animals). While in transect to and from San Andres, two more species not seen elsewhere - the Atlantic brown booby (*Sula leucogaster*) and the roseate tern (*Sterna dougallii*) - were documented although in small numbers.  Off transect, we registered five additional bird species: , laughing gull (*Leucophaeus atricilla*) and great blue heron (*Ardea herodias*) in Serranilla, cattle egret (*Bubulcus ibis*), royal tern (*Thalasseus maximus*) and osprey (*Pandion haliaetus*) in Bajo Nuevo. 

```{r include = T, echo = F}
birds_census %>% 
  group_by(location, common_name) %>% 
  summarize(sum_ind = sum(n_ind),
            n_stations = n_distinct(ps_station_id)) %>% 
  ungroup() %>% 
  mutate(freq_occur = case_when(
    location == "Bajo Nuevo" ~ n_stations/nrow(birds_meta[birds_meta$location == "Bajo Nuevo", ]),
    location == "Serranilla" ~ n_stations/nrow(birds_meta[birds_meta$location == "Serranilla", ]),
    TRUE ~ n_stations/nrow(birds_meta[!birds_meta$location == "Transit", ]))) %>% 
  mutate(freq_occur = round(100*freq_occur, 1)) %>% 
  dplyr::select(-n_stations) %>% 
  pivot_longer(-c(location, common_name)) %>% 
  pivot_wider(names_from = c(location, name), values_from = value) %>% 
  left_join(birds_taxa %>% 
              dplyr::select(common_name, scientific_name, family)) %>% 
  dplyr::select(family, scientific_name, everything(), -common_name) %>% 
  arrange(family, scientific_name) %>% 
  flextable() 
  set_header_labels(values = list(family = "Family", taxa = "Taxa", 
                                  `Bajo Nuevo_sum_ind` = "# individuals",
                                  `Bajo Nuevo_freq_occur` = "% Ocurrence",
                                  `Serranilla_sum_ind` = "# individuals",
                                  `Serranilla_freq_occur` = "% Ocurrence",
                                  `Transit_sum_ind` = "# individuals",
                                  `Transit_freq_occur` = "% Ocurrence")) %>% 
  flextable::add_header_row(values = c("", "", "Bajo Nuevo", "Serranilla", "Transit"), 
                            colwidths = c(1,1,2,2, 2)) %>% 
  merge_v(j = "family") %>% 
  autofit(part = "all") %>%
  flextable::align(align = "center", part = "all")
```

```{r}
readxl::read_xlsx(file.path(ps_exp_path, 
                                        "data/primary/raw",
                                        "birds","Colombia_Seabirds_Fieldsheet_2022_05_17_lat_longs updated.xlsx"), 
                              sheet = "Species Seaflower") %>% 
              janitor::clean_names() %>% 
              dplyr::select(common_name = taxa, binomial) %>% 
              mutate(common_name = str_to_sentence(common_name)) %>% 
  filter(!common_name %in% birds_census$common_name)
```


Metadata


```{r include = T, echo = F}
birds_meta %>% 
  arrange(date) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  dplyr::select(date, location, ps_station_id, lat_start, long_start, time_start, duration_mins) %>% 
  flextable() %>% 
  merge_v(j = "family") %>% 
  autofit(part = "all") %>%
  flextable::align(align = "center", part = "all")
```


