---
title: "Process seabird"
author: 
output:
  word_document:
    toc: no
---

```{r echo = FALSE, include = FALSE, warning = FALSE,prompt = FALSE, error = FALSE, strip.white = TRUE}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, warning = FALSE,
                      prompt = FALSE, error = FALSE, strip.white = TRUE)

knitr::opts_knit$set(progress = TRUE, verbose = TRUE)

options(dplyr.summarise.inform = FALSE)

library(tidyverse)
library(flextable)
library(raster)
library(sf)
library(ggspatial)
library(lubridate)
library(bdc)
library(terra)
source("_prj_vars.R")

knitr::opts_chunk$set(echo = TRUE, fig.cap = TRUE)
  set_flextable_defaults(
  font.family = "Arial", font.size = 9, 
  theme_fun = "theme_vanilla",
  big.mark="", table.layout="autofit")
  
load(file = file.path(ps_exp_path, "data", "secondary", "processed", "admin_spatial_layers.RData"))

edna_meta <- read_csv(file.path(ps_exp_path, 
                               "data", "primary", "processed", "metadata", "edna_meta.csv")) %>% 
  filter(location == "Caribbean")

n_per_loc <- edna_meta %>% 
  group_by(sublocation) %>% 
  summarize(n_stations = n_distinct(ps_station_id),
            n_samples = sum(n_replicates),
            lts = sum(n_replicates*water_quantity_liters))
```

We collected seawater at `r n_per_loc$n_stations[n_per_loc$sublocation == "Serranilla"]` stations in Serranilla and `r n_per_loc$n_stations[n_per_loc$sublocation == "Bajo Nuevo"]` in Bajo Nuevo for e-DNA extraction. At each station, we collected 3 replicates of 2 liter samples both at the surface and at bottom (< `r max(edna_meta$depth_m)` meters) for a total of 130 litres in Bajo Nuevo and 54 in Serranilla. Samples will be processed and analyzed at the Hawaii Institute of Marine Biology and results will be made publicly available. 


```{r include = T, echo = F}
edna_meta %>% 
  group_by(date, sublocation, ps_station_id, lat, lon) %>% 
  summarize(n_samples = sum(n_replicates),
            liters = sum(n_replicates*water_quantity_liters)) %>% 
  mutate_if(is.numeric, round, 3) %>% 
  flextable() %>% 
  merge_v(j = "family") %>% 
  autofit(part = "all") %>%
  flextable::align(align = "center", part = "all")
```

